### 2019-02-06 paper/Paper.tex
Ho sottomesso il pepar

% Please give the surname of the lead author for the running footer
\leadauthor{D'Amico, Gabbolini, Parroni}

\title{Robust background subtraction in traffic environments}
%NM-BC: The title is limited to 10 words (or 90 characters)
\shorttitle{Background subtraction}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[1 \space *]{D'Amico Edoardo}
\author[1 \space *]{Gabbolini Giovanni}
\author[1 \space *]{Parroni Federico}

\affil[1]{Politecnico di Milano, IT}
\affil[*]{These authors contributed equally.}

\maketitle

\begin{abstract}
In this paper we present a method for background subtraction with the aim to work on a 24/24h videos scenario, real time, robust to weather changes and capable to keep foreground objects detected for a large amount of time. The work is the base to implement a monitoring system for dangerous event in the road (possible scenario is the monitoring system on highways). The system can be expanded to detect events such as car driving in the wrong direction, car accidents or people crossing the road.
The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a parametric background modeling paradigm, thus the background is modeled by a history of recently observed pixel
values. The foreground decision depends on a decision
threshold. The background update is based on a learning parameter. Both parameters are extended to dynamic per-pixel state variables and introduce dynamic controllers for each of them. Furthermore, both controllers are steered by an estimate of the background dynamics.
All the hyperparameters of the models have been studied and tuned minutely to accomplish the aimed task.
\end {abstract}

\begin{keywords}
    Background subtraction | PBAS | Car | Traffic | Road | Highway | Surveillance | Stationary camera
\end{keywords}

\subsection*{Introduction}

\subsection*{Related work}

\subsection*{Proposed approach}

\subsection*{Experiments}

\subsection*{Conclusion}

### 2019-02-07 ytdownloader.py
Auto stash before merge of "master" and "origin/master"

@@ -4,6 +4,7 @@ import os
 import math
 # import time
 
+
 #url = "https://www.youtube.com/watch?v=Zwt3ZOej3_U"
 DOWNLOAD_FOLDER = 'dataset'         # download folder
 SAVE_VIDEO = True                   # set to false to save single frames
@@ -14,21 +15,6 @@ VIDEO_CHUNKS_FRAMES = 30*(FPS*60)    # save video at chunks of frames (30 min @ 
 url = "https://www.youtube.com/watch?v=1EiC9bvVGnk"
 video = pafy.new(url)
 print('> Info:')
-print(video)
-
-# best = video.getbest(preftype="mp4")
-print("> Streams:")
-streams = [s for s in video.streams if s.extension == 'mp4']
-index = 0
-for s in streams:
-    print('[{}]'.format(index), s.resolution, s.extension, s.get_filesize())
-    index += 1
-
-print('Choose video resolution: ', end='')
-choice = int(input()[0])
-
-stream = streams[choice]
-videourl = stream.url
 
 filename_max_length = min(20, len(video.title))
 video_folder = "{}/{}".format(DOWNLOAD_FOLDER, video.title[0:filename_max_length])
@@ -61,6 +47,7 @@ ret = True
 frames_count = 0
 video_chunk = 0
 while True:
+
     if SAVE_VIDEO:
         current_frame = frames_count % VIDEO_CHUNKS_FRAMES
         # check if a new chunk must be created

### 2019-02-07 paper/Paper.tex
Related Work

@@
 \subsection*{Related work}
+Over the recent past, a multitude of algorithms and methods for background modeling have been developed.
+One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM)
+???.
+Here, each pixels is modeled as a mixture of weighted Gaussian distributions.
+Pixels, which are detected as background are used to improve the Gaussian mixtures by an iterative update rule.
+A very important non-parametric method is the ViBe ???.
+Each pixel in the background model is defined by a history of the N most recent image values at each pixel and uses a random scheme to update them.
+More over, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another random selection method.The preceding scheme is very similar to the approach followed by the PBAS algorithm ???.
+It can be categorized as a non-parametric method, since it  uses a history of N image values as the background model, and uses a random update rule similar to the one used by the ViBe algorithm.
+However, in Vibe, the randomness parameters as well as the decision threshold are fixed for all pixels.
+In contrast, in the PBAS algorithm these values are not treated as parameters, but instead as adaptive state variables, which can dynamically change over time for each pixel separately.
 

### 2019-02-07 paper/Paper.tex
Intro

--- 2019-02-06 paper/Paper.tex
@@
 \begin{abstract}
-In this paper we present a method for background subtraction with the aim to work on a 24/24h videos scenario, real time, robust to weather changes and capable to keep foreground objects detected for a large amount of time.
-The work is the base to implement a monitoring system for dangerous event in the road (possible scenario is the monitoring system on highways).
-The system can be expanded to detect events such as car driving in the wrong direction, car accidents or people crossing the road.
-The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a parametric background modeling paradigm, thus the background is modeled by a history of recently observed pixel
-values. The foreground decision depends on a decision
-threshold.
-The background update is based on a learning parameter.
-Both parameters are extended to dynamic per-pixel state variables and introduce dynamic controllers for each of them.
-Furthermore, both controllers are steered by an estimate of the background dynamics.
-All the hyperparameters of the models have been studied and tuned minutely to accomplish the aimed task.
+In this paper we present a method for background subtraction with the aim to work on a 24/24h videos
+scenario, real time, robust to weather changes and capable to keep foreground objects detected for a
+large amount of time. The work is the base to implement a monitoring system for dangerous event in
+the road (possible scenario is the monitoring system on highways). The system can be expanded to
+detect events such as car driving in the wrong direction, car accidents or people crossing the road.
+The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a parametric background
+modeling paradigm, thus the background is modeled by a history of recently observed pixel values.
+The foreground decision depends on a decision threshold. The background update is based on a learning
+parameter. Both parameters are extended to dynamic per-pixel state variables and introduce dynamic
+controllers for each of them. Furthermore, both controllers are steered by an estimate of the background
+dynamics. All the hyperparameters of the models have been studied and tuned minutely to accomplish the
+aimed task.
 \end {abstract}
@@
 \subsection*{Introduction}
+Background subtraction and foreground detection are the basic tasks of many real application systems,
+for example, surveillance systems, autonomous vehicles, semantic image analysis. Background subtraction
+consists in finding a category for every pixel in a single or (like in this case) in a frames sequence
+of a video and saying if that pixels are part of background or foreground. At the end, the algorithm take as
+as input a video and output a binary mask, where 1s are foreground pixels and 0 background pixels.
+We restricted our domain to traffic monitoring, in particular this work should put the basis for an automated
+tool to identify cars and detect anomalies in highways and roads 24/24h (e.g. car accidents,
+traffic jam, wrong-direction driving) using a standard RGB stationary camera.
+Critical aspects that must be taken into account while
+developing such algorithm are a lot, first of all, variable and bad weather conditions and illumination
+changes. In fact, it is very hard to distinguish moving objects in presence of heavy rain, fog or snow.
+Also during night the environmental situation varies a lot from the day and our algorithm has to adapt
+continuosly due to this facts. Other problems that we faced are shadows and intermittent object motion.
+From the majority of the current algorithms, shadows are treated as foreground because they move along with
+objects; a little number of them adopts a 3-class classification (background, foreground, shadow). We
+decided to provide a binary classification, considering shadows as part of the background. With regards to
+the latter problem, here the challenge is to correctly detect an initially stationary object that begins to
+move or when an object that was static starts moving again. We fine-tuned some parameters to make sure
+that a moving car that stops after a while will be classified as foreground for a long time before turning
+into background, while the opposite event (from static to moving) is easier to handle.
+The entire algorithm has been implemented in C++, since compiled code performances are fundamental to
+allow the system to work in real-time. In \cite{pbasandsceneanalysisfpga}, we can find an integration of the basic algorithm in
+a FPGA device, which exploit and hardware implementation to reduce a lot the computational time.
 

### 2019-02-07 ytdownloader.py
Insert video url. Small graphical changes

@@ -11,7 +11,10 @@ FPS = 30
 VIDEO_CODEC = cv2.VideoWriter_fourcc(*'mp4v')
 VIDEO_CHUNKS_FRAMES = 30*(FPS*60)    # save video at chunks of frames (30 min @ 25 fps)
 
-url = "https://www.youtube.com/watch?v=1EiC9bvVGnk"
+# url = "https://www.youtube.com/watch?v=1EiC9bvVGnk"
+# url = 'https://www.youtube.com/watch?v=ZTa4ap3i278'
+print('Insert youtube video url: ')
+url = input()
 video = pafy.new(url)
 print('> Info:')
 print(video)
@@ -25,7 +28,11 @@ for s in streams:
     index += 1
 
 print('Choose video resolution: ', end='')
-choice = int(input()[0])
+try:
+    choice = int(input()[0])
+except Exception as exc:
+    print('Wrong input, exiting...')
+    exit(1)
 
 stream = streams[choice]
 videourl = stream.url
@@ -41,7 +48,7 @@ def get_frame(capture):
     return ret, frame
 
 def print_progress(current, tot, prefix=''):
-    scale = 30 / tot
+    scale = 60 / tot
     left = math.floor((tot-current) * scale)
     curr = math.ceil(current * scale)
     print('{}: |={}{}|'.format(prefix, '='*curr, ' '*left), end='\r')

### 2019-02-07 paper/Bibliography.tex
bibliography added

\begin{thebibliography}{9}

\bibitem{gmm} 
C. Stauffer and W. Grimson. Adaptive background mixture
models for real-time tracking. In Computer Vision and Pattern Recognition, 1999. IEEE Computer Society Conference on., volume 2, pages 2 vol. (xxiii+637+663), 1999.

\end{thebibliography}

### 2019-02-07 paper/Paper.tex
bibliography added
citation added

@@
 One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM)
-???.
+\cite{gmm}.
 Here, each pixels is modeled as a mixture of weighted Gaussian distributions.
 Pixels, which are detected as background are used to improve the Gaussian mixtures by an iterative update rule.
-A very important non-parametric method is the ViBe ???.
+A very important non-parametric method is the ViBe \cite{vibe}.
 Each pixel in the background model is defined by a history of the N most recent image values at each pixel and uses a random scheme to update them.

### 2019-02-07 readvideo.cpp
menu for saving frames

@@ -1,12 +1,96 @@
 #include <iostream>
 #include <string>
 #include <opencv2/opencv.hpp>
+#include <sys/stat.h> 
 #include "pbas.cpp"
 
 using namespace std;
 using namespace cv;
 
+int state;
+int starting_minute;
+bool save_frames;
+vector<tuple<int, int>> time_slots;
+int capture_every;
+string folder_name;
+
+void go_to(int s){
+    switch (s) {
+        case 0:
+            state = 0;
+            cout << "want to save frames? (y/n)" << endl;
+            break;
+        case 1:
+            state = 1;
+            cout << "put starting minute in which to save (x) or enter to go ahead" << endl;
+            break;
+        case 2:
+            state = 2;
+            cout << "put ending minute in which stop to save (x)" << endl;
+            break;
+        case 3:
+            state = 3;
+            cout << "save a frame every? (x, in seconds). enter for default value 1" << endl;
+            break;
+        case 4:
+            state = 4;
+            cout << "select a folder name in which to save (enter for a default name)" << endl;
+            break;
+    }
+}
+
+int process(string & line){
+    switch (state)
+    {
+        case 0:
+            if(line == ""){
+                go_to(0);
+            }
+            if(line == "y"){
+                go_to(1);
+                save_frames = true;
+            }
+            else{
+                return -1;
+            }
+            break;
+        case 1:
+            if(line == ""){
+                go_to(3);
+            }
+            else{
+                starting_minute = stoi(line);
+                go_to(2);
+            }
+            break;
+        case 2:
+            if(line == ""){
+                go_to(2);
+            }
+            else{
+                time_slots.push_back(make_tuple(starting_minute, stoi(line)));
+                go_to(1);
+            }
+            break;
+        case 3:
+            if(line == ""){
+                capture_every = 1;
+                go_to(4);
+            }
+            else{
+                capture_every = stoi(line);
+                go_to(4);
+            }
+            break;
+        case 4:
+            folder_name = line;
+            return -1;
+            break;
+    }
+}
+
 int main(int argc, char const *argv[]) {
+    // init
[…]

### 2019-02-07 readvideo.cpp
Auto stash before merge of "master" and "origin/master"

@@ -167,7 +167,7 @@ int main(int argc, char const *argv[]) {
         if (save_frames) {
             act_seconds += 1 / fps;
             cout << act_seconds << endl;
-            if (act_seconds >= get<0>(time_slots[count_time_slot]) && act_seconds <= get<1>(time_slots[count_time_slot])){
+            if (act_seconds >= get<0>(time_slots[count_time_slot])*60 && act_seconds <= get<1>(time_slots[count_time_slot])*60){
                 partial_seconds += 1 / fps;
                 if(partial_seconds >= capture_every){
                     cout << "capturing frame" << endl;
@@ -178,7 +178,7 @@ int main(int argc, char const *argv[]) {
                     partial_seconds = 0;
                 }
             }
-            else if (act_seconds > get<1>(time_slots[count_time_slot]) && count_time_slot < time_slots.size() - 1){
+            else if (act_seconds > get<1>(time_slots[count_time_slot])*60 && count_time_slot < time_slots.size() - 1){
                 count_time_slot++;
                 act_subfolder_base = folder_base + "/" + to_string(get<0>(time_slots[count_time_slot])) + "_" + to_string(get<1>(time_slots[count_time_slot]));
                 mkdir(act_subfolder_base.c_str(), 0777);

### 2019-02-07 paper/Paper.tex
PBAS section completed

@@
 \begin{abstract}
-In this paper we present a method for background subtraction with the aim to work on a 24/24h videos
+\textit{	Abstract-} In this paper we present a method for background subtraction with the aim to work on a 24/24h videos
 scenario, real time, robust to weather changes and capable to keep foreground objects detected for a
@@
 \subsection*{Proposed approach}
+To solve the presented problem we have studied deeply the PBAS algorithm, how it works, and how the hyperparameters could change the behaviour of it.
+After the studying of the algorithm has been chosen the parameters that have achived the better performance in our scenario.
+At the end have been studied how to enhanche the performance of the algorithm solving the issue of the presence of the shadows in the video, using techniques of shadows removal.
+\subsubsection{PBAS algorithm}
 
+\begin{figure*}
+  \includegraphics[width=\textwidth]{Figures/PBAS_state_machine.png}
+  \caption{PBAS state machine.}
+\end{figure*}
+
+This section describes the Pixel-Based Adaptive Segmenter, which follows a non-parametric paradigm. Thus,
+every pixel x i is modeled by an array of recently observed background values.
+The method consists of several components which are depicted as a state machine in Figure 1.
+As a central component, the decision block decides for or against foreground
+based on the current image and a background model $B(x_i)$.
+This decision is based on the per-pixel threshold $R(x_i)$.
+Moreover, the background model has to be updated over time in order to allow for gradual background changes.
+In the model, this update depends on a per-pixel learning parameter $T(x_i)$.
+\newline The background model $B(x_i)$ is defined by an array of N recently observed pixel values:
+\begin{equation}
+\mathbf{B}(x_i) = \{B_1(x_i),...,B_k(x_i),...,B_N(x_i)\}
+\end{equation}
+A pixel $x_i$ is decided to belong to the background, if its
+pixel value $I(x_i) $ is closer than a certain decision threshold
+$R(x_i)$ to at least $K$ of the $N$ background values. Thus,
+the foreground segmentation mask is calculated as:
+\begin{equation}
+F(x_i) = \begin{cases} 1,& \#\{dist(I(x_i), B_k(x_i)) < R(x_i)\} < K \\0 &else\end{cases}
+\end{equation}
+$F = 1$ implies foreground.
+The decision involves two parameters: (1) The distance threshold $R(x_i)$,which is defined for each pixel separately and which can change dynamically; and (2) the minimum number $K$, which is a fixed global parameter.\newline
+The background model is only updated for those pixels that
+are currently background (i.e. $F(x_i) = 0$).
+Updating means that for a certain index $k \in \{1...N \}$(chosen uniformly at random), the corresponding background model
+value $B_k(x_i)$ is replaced by the current pixel value $I(x_i)$.
+This allows the current pixel value to be ”learned” into
+the background model. This update, however, is only performed with probability $p = 1/T(x_i)$.Otherwise no update
+is carried out at all. Therefore, the parameter $T(x_i)$ defines
+the update rate. The higher $T(x_i)$ the less likely a pixel
+will be updated.
+Is also update (with probability $p = 1/T(x_i)$) a randomly chosen neighboring pixel $y i \in N(x_i)$. Thus, the
+background model $B_k(y_i)$ at this neighboring pixel is replaced by its current pixel value $V(y_i)$. \newline
+To change the decision treshold $R(x_i)$ is saved an array of recently observed pixel values in the background model $B(x_i)$, we also create an array $\mathbf{D}(x_i) = \{D_1(x_i ),...,D_N(x_i)\}$ of minimal decision distances.
+Whenever an update of $B_k(x_i)$ is carried out, the currently observed minimal distance $d_{min}(x_i) = min_k dist(I(x_i), B_k(x_i))$ is written to this array: $D_k(x_i) \leftarrow d_{min}(x_i)$.
+Thus, a history of minimal decision distances is created.
+The average of these values $\overline{d}_{min}(x_i) = \frac{1}{N} \sum_k D_k(x_i)$ is a measure of the background dynamics.
+Thanks to that
+the decision threshold can be dynamically adapted as follows:
+\begin{equation}
+R(x_i)=\begin{cases} R(x_i)\cdot(1-R{inc/dec}) &if\ R(x_i)> \overline{d}_{min}(x_i)\cdot R_{scale} \\ R(x_i)\cdot (1+R{inc/dec}) &else  \end{cases}
+\end{equation}
+Here, $R_{inc/dec}$, $R_{scale}$ are fixed parameters. This can be seen as a dynamic controller for the state variable $R(x_i)$.
+\newline
+Independent of the foreground state $F(x_i)$, every object will be merged into the background depending on the learning parameter
+$T(x_i)$. To alleviate the problem, the idea is to introduce a
+(second) dynamic controller for $T(x_i)$ following the assumption that pixels are mostly wrongly classified as foreground in areas of high dynamic background the strength of the adjustment in the controller can be adapted using the dynamic estimator $\overline{d}_{min}(x_i)$.
+So can be defined:
+\begin{equation}
+T(x_i)=\begin{cases} T(x_i)+\frac{T_{inc}}{\overline{d}_{min}(x_i)} &if \ F(x_i)=1 \\ T(x_i)-\frac{T_{inc}}{\overline{d}_{min}(x_i)} &if \ F(x_i)=0 \end{cases}
+\end{equation}
+$T_{inc}$ and $T_{dec}$ are fixed parameters. Are also defined $T_{lower}$ and $T_{upper}$ such that the values can not go out of a specific bound. 
 \subsection*{Experiments}

### 2019-02-07 paper/Bibliography.tex
citation added
Added section Experiments. Images to fix.

@@
 C. Stauffer and W. Grimson. Adaptive background mixture
-models for real-time tracking. In Computer Vision and Pattern Recognition, 1999. IEEE Computer Society Conference on., volume 2, pages 2 vol. (xxiii+637+663), 1999.
+models for real-time tracking. In Computer Vision and Pattern Recognition, 1999.
+IEEE Computer Society Conference on., volume 2, pages 2 vol. (xxiii+637+663), 1999.
+
+\bibitem{vibe}
+O. Barnich and M. Van Droogenbroeck. Vibe: A universal
+background subtraction algorithm for video sequences. Im-
+age Processing, IEEE Transactions on, 20(6):1709 –1724,
+june 2011.
+
+\bibitem{pbas_and_scene_analysis_fpga}
+T. Kryjak, M. Komorkiewicz and M. Gorgon.
+Real-time Foreground Object Detection Combining
+the PBAS Background Modelling Algorithm
+and Feedback from Scene Analysis Module.
+INTL Journal of Electronics and Telecommunications, 2014, VOL. 60, NO. 1, PP. 61–72
 

### 2019-02-07 paper/Paper.tex
Added section Experiments. Images to fix.

--- 2019-02-07 paper/Paper.tex
@@
 
+\section*{Abstract}
 \begin{abstract}
@@
 \begin{keywords}
-    Background subtraction | PBAS | Car | Traffic | Road | Highway | Surveillance | Stationary camera
+    Background subtraction | Foreground detection | PBAS algorithm | Car | Traffic | Road | Highway | Surveillance |
+    Stationary camera | Image processing and analysis
 \end{keywords}
 
-\subsection*{Introduction}
+\section*{Introduction}
 Background subtraction and foreground detection are the basic tasks of many real application systems,
@@
 traffic jam, wrong-direction driving) using a standard RGB stationary camera.
-Critical aspects that must be taken into account while
-developing such algorithm are a lot, first of all, variable and bad weather conditions and illumination
-changes. In fact, it is very hard to distinguish moving objects in presence of heavy rain, fog or snow.
-Also during night the environmental situation varies a lot from the day and our algorithm has to adapt
-continuosly due to this facts. Other problems that we faced are shadows and intermittent object motion.
+Critical aspects that must be taken into account while developing such algorithm are a lot, first of all,
+variable and bad weather conditions and illumination changes. In fact, it is very hard to distinguish
+moving objects in presence of heavy rain, fog or snow. Also during night the environmental situation varies
+a lot from the day and our algorithm has to adapt continuosly due to this facts. 
+Other problems that we faced are shadows and intermittent object motion.
 From the majority of the current algorithms, shadows are treated as foreground because they move along with
@@
 The entire algorithm has been implemented in C++, since compiled code performances are fundamental to
-allow the system to work in real-time. In \cite{pbasandsceneanalysisfpga}, we can find an integration of the basic algorithm in
-a FPGA device, which exploit and hardware implementation to reduce a lot the computational time.
+allow the system to work in real-time. In \cite{pbas_and_scene_analysis_fpga}, we can find an integration of
+the basic algorithm in a FPGA device, which exploit and hardware implementation to reduce a lot the
+computational time.
 
-\subsection*{Related work}
+\section*{Related work}
 Over the recent past, a multitude of algorithms and methods for background modeling have been developed.
 One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM)
-\cite{gmm}.
-Here, each pixels is modeled as a mixture of weighted Gaussian distributions.
-Pixels, which are detected as background are used to improve the Gaussian mixtures by an iterative update rule.
-A very important non-parametric method is the ViBe \cite{vibe}.
-Each pixel in the background model is defined by a history of the N most recent image values at each pixel and uses a random scheme to update them.
-More over, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another random selection method.The preceding scheme is very similar to the approach followed by the PBAS algorithm ???.
-It can be categorized as a non-parametric method, since it  uses a history of N image values as the background model, and uses a random update rule similar to the one used by the ViBe algorithm.
-However, in Vibe, the randomness parameters as well as the decision threshold are fixed for all pixels.
-In contrast, in the PBAS algorithm these values are not treated as parameters, but instead as adaptive state variables, which can dynamically change over time for each pixel separately.
+\cite{gmm}. Here, each pixels is modeled as a mixture of weighted Gaussian distributions. Pixels, which
+are detected as background are used to improve the Gaussian mixtures by an iterative update rule. A very
+important non-parametric method is the ViBe \cite{vibe}. Each pixel in the background model is defined by
+a history of the N most recent image values at each pixel and uses a random scheme to update them.
+More over, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another
+random selection method. The preceding scheme is very similar to the approach followed by the PBAS algorithm.
+It can be categorized as a non-parametric method, since it uses a history of N image values as the background
+model, and uses a random update rule similar to the one used by the ViBe algorithm. However, in Vibe, the
+randomness parameters as well as the decision threshold are fixed for all pixels. In contrast, in the PBAS
+algorithm these values are not treated as parameters, but instead as adaptive state variables, which can
+dynamically change over time for each pixel separately.
 
-\subsection*{Proposed approach}
+\section*{Proposed approach}
 
-\subsection*{Experiments}
+\section*{Experiments}
+In this section we first report the results obtained by varying the main model parameters. We also include the
+adopted settings that perform best in our scenario. Then, we show some real-case usages of the algorithm,
+comparing the different environmental conditions that we found in 24/24h live videos.
 
-\subsection*{Conclusion}
+\subsection*{N}
+$N$ represents how many matrices are kept in memory to model the background.
+Each of them stores the history of pixels values.
+The greater, the more complex background can be handled but at the cost of heavier
+computation. At $N=35$, experimental results show that performance saturates.
+\begin{figure*}[!t]
+\centering
+\subfloat[][N = 2]{
+    \includegraphics[width=0.32\textwidth]{Figures/N2.jpg}
+    %\label{fig:N2}
+}
+\subfloat[][N = 10]{
+    \includegraphics[width=0.32\textwidth]{Figures/N10.jpg}
+}
+\subfloat[][N = 20]{
+    \includegraphics[width=0.32\textwidth]{Figures/N20.jpg}
+}
+\end{figure*}
+We can see the poor mask of the cars on the left road when $N=2$, and the
+accurate mask of the lines on the front right when $N=20$.
+
+\subsection*{K}
+$K$ tells the number of samples from $B$ that have to be closer than $R$ in order to
+classify the pixel as background. The higher, the more difficult for a pixel to be
+classified as background.
+\begin{figure*}[!t]
+\centering
+\subfloat[][K = 1]{
+    \includegraphics[width=0.32\textwidth]{Figures/K1.jpg}
+}
+\subfloat[][K = 2]{
+    \includegraphics[width=0.32\textwidth]{Figures/K2.jpg}
+}
+\subfloat[][K = 4]{
+    \includegraphics[width=0.32\textwidth]{Figures/K4.jpg}
+}
+\end{figure*}
+We can notice that setting $K$ too high brings to classify many pixels as foreground.
+
+\subsection*{T max}
+$T_max$ controls the lowest probability of updating a pixel and putting it the
+background model $B$.
+
+\subsection*{R}
+Eq ?????????? is similar to the equation of a pure proportional controller.
+In this case the $R(x,y)$ has to follow $d_minavg(x,y) * R_scale$. So:
+\begin{itemize}
+    \item $R_{incdec}$: setting too high lead to oscillation around the target and if
+    set too high leads to divergence. The default value is 0.05 which leads to fast
+    enough adaptation to the target value (in more or less 30 frames the target is
+    reached). Assuming that the backgroud isnt changing much there is no reason to
+    raise it.
+    \item $R_{scale}$: tells how much the algorithm is sensitive, in fact it is a direct
+    control of the threshold. Setting too high it leads to really few false positives
+    but also decreases true positives; setting too low, will increase false positives.
+    \item $R_{lower}$: sets a limit on how much $R$ can decrease its value. Setting too
+    low, it will lead to a lot of false positives in area that are foreground,
+    especially if the video is noise, so especially is zones that are clearly
+    background are changing their values of high amounts due to noise. Setting too
+    high it will reduce true positives. So the choice of it should be lead by the
+    video quality, more than the application domain.
+\end{itemize}
+\begin{figure*}[!t]
+    \centering
+    \subfloat[][$R_{scale}$ = 5]{
+        \includegraphics[width=0.32\textwidth]{Figures/R_scale5.jpg}
+    }
+    \subfloat[][$R_{scale}$ = 15]{
+        \includegraphics[width=0.32\textwidth]{Figures/R_scale15.jpg}
+    }
+    \subfloat[][$R_{scale}$ = 50]{
+        \includegraphics[width=0.32\textwidth]{Figures/R_scale50.jpg}
+    }
+\end{figure*}
+As we see, as $R_{scale}$ grows the tree leaves that are waved by the wind are not detected
+anymore as foreground but also a part of the car is lost.
+
+\subsection*{alpha}
+From eq ?????????? we see that alpha tells how much to weight the image gradients
+with respect to the image values in the distance computation.
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/gradients.jpg}
+    \caption{Gradients}
+    \label{fig:gradients}
+\end{figure}
+As we see, the image gradients give a good cue to distinguish the street from the cars
+(image absolute values are not, since many cars have almost the same color of the
+street). So alpha should be set high, in order to take into consideration this fact.
+But notice that setting too high will lead to noisy masks, especially with noisy
+videos that have small variations of the gradient from frame to frame.
+\begin{figure*}[!t]
+    \centering
+    \subfloat[][$\alpha$ = 0]{
+        \includegraphics[width=0.32\textwidth]{Figures/alpha0.jpg}
+    }
+    \subfloat[][$\alpha$ = 10]{
+        \includegraphics[width=0.32\textwidth]{Figures/alpha10.jpg}
+    }
+    \subfloat[][$\alpha$ = 20]{
+        \includegraphics[width=0.32\textwidth]{Figures/alpha20.jpg}
+    }
+    \subfloat[][$\alpha$ = 50]{
+        \includegraphics[width=0.32\textwidth]{Figures/alpha50.jpg}
+    }
+\end{figure*}
+
+\section*{Conclusion}

### 2019-02-08 paper/Paper.tex
Added text as caption to some figures

@@
 One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM)
+
 \cite{gmm}. Here, each pixels is modeled as a mixture of weighted Gaussian distributions. Pixels, which
@@
 \section*{Proposed approach}
+To solve the presented problem we have studied deeply the PBAS algorithm, how it works, and how the hyperparameters could change the behaviour of it.
+After the studying of the algorithm has been chosen the parameters that have achived the better performance in our scenario.
+At the end have been studied how to enhanche the performance of the algorithm solving the issue of the presence of the shadows in the video, using techniques of shadows removal.
+\subsubsection{PBAS algorithm}
+
+\begin{figure*}
+  \includegraphics[width=\textwidth]{Figures/PBAS_state_machine.png}
+  \caption{PBAS state machine.}
+\end{figure*}
+
+This section describes the Pixel-Based Adaptive Segmenter, which follows a non-parametric paradigm. Thus,
+every pixel x i is modeled by an array of recently observed background values.
+The method consists of several components which are depicted as a state machine in Figure 1.
+As a central component, the decision block decides for or against foreground
+based on the current image and a background model $B(x_i)$.
+This decision is based on the per-pixel threshold $R(x_i)$.
+Moreover, the background model has to be updated over time in order to allow for gradual background changes.
+In the model, this update depends on a per-pixel learning parameter $T(x_i)$.
+\newline The background model $B(x_i)$ is defined by an array of N recently observed pixel values:
+\begin{equation}
+\mathbf{B}(x_i) = \{B_1(x_i),...,B_k(x_i),...,B_N(x_i)\}
+\end{equation}
+A pixel $x_i$ is decided to belong to the background, if its
+pixel value $I(x_i) $ is closer than a certain decision threshold
+$R(x_i)$ to at least $K$ of the $N$ background values. Thus,
+the foreground segmentation mask is calculated as:
+\begin{equation}
+F(x_i) = \begin{cases} 1,& \#\{dist(I(x_i), B_k(x_i)) < R(x_i)\} < K \\0 &else\end{cases}
+\end{equation}
+$F = 1$ implies foreground.
+The decision involves two parameters: (1) The distance threshold $R(x_i)$,which is defined for each pixel separately and which can change dynamically; and (2) the minimum number $K$, which is a fixed global parameter.\newline
+The background model is only updated for those pixels that
+are currently background (i.e. $F(x_i) = 0$).
+Updating means that for a certain index $k \in \{1...N \}$(chosen uniformly at random), the corresponding background model
+value $B_k(x_i)$ is replaced by the current pixel value $I(x_i)$.
+This allows the current pixel value to be ”learned” into
+the background model. This update, however, is only performed with probability $p = 1/T(x_i)$.Otherwise no update
+is carried out at all. Therefore, the parameter $T(x_i)$ defines
+the update rate. The higher $T(x_i)$ the less likely a pixel
+will be updated.
+Is also update (with probability $p = 1/T(x_i)$) a randomly chosen neighboring pixel $y i \in N(x_i)$. Thus, the
+background model $B_k(y_i)$ at this neighboring pixel is replaced by its current pixel value $V(y_i)$. \newline
+To change the decision treshold $R(x_i)$ is saved an array of recently observed pixel values in the background model $B(x_i)$, we also create an array $\mathbf{D}(x_i) = \{D_1(x_i ),...,D_N(x_i)\}$ of minimal decision distances.
+Whenever an update of $B_k(x_i)$ is carried out, the currently observed minimal distance $d_{min}(x_i) = min_k dist(I(x_i), B_k(x_i))$ is written to this array: $D_k(x_i) \leftarrow d_{min}(x_i)$.
+Thus, a history of minimal decision distances is created.
+The average of these values $\overline{d}_{min}(x_i) = \frac{1}{N} \sum_k D_k(x_i)$ is a measure of the background dynamics.
+Thanks to that
+the decision threshold can be dynamically adapted as follows:
+\begin{equation}
+R(x_i)=\begin{cases} R(x_i)\cdot(1-R{inc/dec}) &if\ R(x_i)> \overline{d}_{min}(x_i)\cdot R_{scale} \\ R(x_i)\cdot (1+R{inc/dec}) &else  \end{cases}
+\end{equation}
+Here, $R_{inc/dec}$, $R_{scale}$ are fixed parameters. This can be seen as a dynamic controller for the state variable $R(x_i)$.
+\newline
+Independent of the foreground state $F(x_i)$, every object will be merged into the background depending on the learning parameter
+$T(x_i)$. To alleviate the problem, the idea is to introduce a
+(second) dynamic controller for $T(x_i)$ following the assumption that pixels are mostly wrongly classified as foreground in areas of high dynamic background the strength of the adjustment in the controller can be adapted using the dynamic estimator $\overline{d}_{min}(x_i)$.
+So can be defined:
+\begin{equation}
+T(x_i)=\begin{cases} T(x_i)+\frac{T_{inc}}{\overline{d}_{min}(x_i)} &if \ F(x_i)=1 \\ T(x_i)-\frac{T_{inc}}{\overline{d}_{min}(x_i)} &if \ F(x_i)=0 \end{cases}
+\end{equation}
+$T_{inc}$ and $T_{dec}$ are fixed parameters. Are also defined $T_{lower}$ and $T_{upper}$ such that the values can not go out of a specific bound.
 
@@
 }
+\caption{We can see the poor mask of the cars on the left road when $N=2$, and the
+accurate mask of the lines on the front right when $N=20$.}
 \end{figure*}
-We can see the poor mask of the cars on the left road when $N=2$, and the
-accurate mask of the lines on the front right when $N=20$.
 
@@
 }
+\caption{Setting $K$ too high brings to classify many pixels as foreground, with the risk of false positives.}
 \end{figure*}
-We can notice that setting $K$ too high brings to classify many pixels as foreground.
 
@@
 
-\section*{Conclusion}
+\subsection*{Conclusion}

### 2019-02-08 readvideo.cpp
new default values

@@ -3,6 +3,7 @@
 #include <opencv2/opencv.hpp>
 #include <sys/stat.h> 
 #include "pbas.cpp"
+#include <limits>
 
 using namespace std;
 using namespace cv;
@@ -22,19 +23,19 @@ void go_to(int s){
             break;
         case 1:
             state = 1;
-            cout << "put starting minute in which to save (x) or enter to go ahead" << endl;
+            cout << "put minute in which start saving or enter to go ahead" << endl;
             break;
         case 2:
             state = 2;
-            cout << "put ending minute in which stop to save (x)" << endl;
+            cout << "put minute in which to stop saving or enter to record till the end" << endl;
             break;
         case 3:
             state = 3;
-            cout << "save a frame every? (x, in seconds). enter for default value 1" << endl;
+            cout << "save a frame every? (x, in seconds). enter for default value 60" << endl;
             break;
         case 4:
             state = 4;
-            cout << "select a folder name in which to save (enter for a default name)" << endl;
+            cout << "folder name where to save: (enter for a default name)" << endl;
             break;
     }
 }
@@ -65,7 +66,8 @@ int process(string & line){
             break;
         case 2:
             if(line == ""){
-                go_to(2);
+                time_slots.push_back(make_tuple(starting_minute, numeric_limits<int>::max()/60));
+                go_to(1);
             }
             else{
                 time_slots.push_back(make_tuple(starting_minute, stoi(line)));
@@ -74,7 +76,7 @@ int process(string & line){
             break;
         case 3:
             if(line == ""){
-                capture_every = 1;
+                capture_every = 60;
                 go_to(4);
             }
             else{
@@ -148,6 +150,7 @@ int main(int argc, char const *argv[]) {
         mkdir(act_subfolder_base.c_str(), 0777);
     }
 
+    // process
     while(1) {
         cap >> frame;
         if (frame.empty()) break;
@@ -184,7 +187,6 @@ int main(int argc, char const *argv[]) {
                 mkdir(act_subfolder_base.c_str(), 0777);
             }
         }
-
 
         // imshow("Shadows CNCC", pbas->shadow_cncc);
         // moveWindow("Shadows CNCC", 750,20);

### 2019-02-08 paper/Paper.tex

@@
 }
-\caption{We can see the poor mask of the cars on the left road when $N=2$, and the
-accurate mask of the lines on the front right when $N=20$.}
 \end{figure*}
+We can see the poor mask of the cars on the left road when $N=2$, and the
+accurate mask of the lines on the front right when $N=20$.
 
@@
 }
-\caption{Setting $K$ too high brings to classify many pixels as foreground, with the risk of false positives.}
 \end{figure*}
+We can notice that setting $K$ too high brings to classify many pixels as foreground.
 

### 2019-02-08 paper/Paper.tex
Correct figures positioning

--- 2019-02-08 paper/Paper.tex
@@
 computation. At $N=35$, experimental results show that performance saturates.
-\begin{figure*}[!t]
+\begin{figure}[!t]
 \centering
 \subfloat[][N = 2]{
-    \includegraphics[width=0.32\textwidth]{Figures/N2.jpg}
+    \includegraphics[width=0.45\textwidth]{Figures/N2.jpg}
     %\label{fig:N2}
 }
+\newline
 \subfloat[][N = 10]{
-    \includegraphics[width=0.32\textwidth]{Figures/N10.jpg}
+    \includegraphics[width=0.45\textwidth]{Figures/N10.jpg}
 }
+\newline
 \subfloat[][N = 20]{
-    \includegraphics[width=0.32\textwidth]{Figures/N20.jpg}
+    \includegraphics[width=0.45\textwidth]{Figures/N20.jpg}
 }
@@
 accurate mask of the lines on the front right when $N=20$.}
-\end{figure*}
+\end{figure}
 
@@
 classified as background.
-\begin{figure*}[!t]
+\begin{figure}[!t]
 \centering
 \subfloat[][K = 1]{
-    \includegraphics[width=0.32\textwidth]{Figures/K1.jpg}
+    \includegraphics[width=0.45\textwidth]{Figures/K1.jpg}
 }
+\newline
 \subfloat[][K = 2]{
-    \includegraphics[width=0.32\textwidth]{Figures/K2.jpg}
+    \includegraphics[width=0.45\textwidth]{Figures/K2.jpg}
 }
+\newline
 \subfloat[][K = 4]{
-    \includegraphics[width=0.32\textwidth]{Figures/K4.jpg}
+    \includegraphics[width=0.45\textwidth]{Figures/K4.jpg}
 }
 \caption{Setting $K$ too high brings to classify many pixels as foreground, with the risk of false positives.}
-\end{figure*}
+\end{figure}
 
 \subsection*{T max}
-$T_max$ controls the lowest probability of updating a pixel and putting it the
+$T_{max}$ controls the lowest probability of updating a pixel and putting it the
 background model $B$.
@@
 Eq ?????????? is similar to the equation of a pure proportional controller.
-In this case the $R(x,y)$ has to follow $d_minavg(x,y) * R_scale$. So:
+In this case the $R(x,y)$ has to follow $d_{minavg}(x,y) * R_{scale}$. So:
 \begin{itemize}
@@
 \end{itemize}
-\begin{figure*}[!t]
+\begin{figure}[!t]
     \centering
     \subfloat[][$R_{scale}$ = 5]{
-        \includegraphics[width=0.32\textwidth]{Figures/R_scale5.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/R_scale5.jpg}
     }
+    \newline
     \subfloat[][$R_{scale}$ = 15]{
-        \includegraphics[width=0.32\textwidth]{Figures/R_scale15.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/R_scale15.jpg}
     }
+    \newline
     \subfloat[][$R_{scale}$ = 50]{
-        \includegraphics[width=0.32\textwidth]{Figures/R_scale50.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/R_scale50.jpg}
     }
-\end{figure*}
+\end{figure}
 As we see, as $R_{scale}$ grows the tree leaves that are waved by the wind are not detected
@@
 videos that have small variations of the gradient from frame to frame.
-\begin{figure*}[!t]
+\begin{figure}[!t]
     \centering
     \subfloat[][$\alpha$ = 0]{
-        \includegraphics[width=0.32\textwidth]{Figures/alpha0.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/alpha0.jpg}
     }
+    \newline
     \subfloat[][$\alpha$ = 10]{
-        \includegraphics[width=0.32\textwidth]{Figures/alpha10.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/alpha10.jpg}
     }
+    \newline
     \subfloat[][$\alpha$ = 20]{
-        \includegraphics[width=0.32\textwidth]{Figures/alpha20.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/alpha20.jpg}
     }
+    \newline
     \subfloat[][$\alpha$ = 50]{
-        \includegraphics[width=0.32\textwidth]{Figures/alpha50.jpg}
+        \includegraphics[width=0.45\textwidth]{Figures/alpha50.jpg}
     }
-\end{figure*}
+\end{figure}
 

### 2019-02-08 comparevideos.cpp
Fixed TOPI MORTI (in theory)

@@ -1,0 +1,88 @@
+#include <iostream>
+#include <string>
+#include <opencv2/opencv.hpp>
+#include <sys/stat.h> 
+#include "pbas.cpp"
+//#include <limits>
+
+using namespace std;
+using namespace cv;
+
+int current_frame = 0;
+float FPS = 30;
+
+int main(int argc, char const *argv[]) {
+    // init
+    int N = 30;
+    int K = 3;
+    float R_incdec = 0.05;
+    int R_lower = 18;
+    int R_scale = 5;
+    float T_dec = 0.05;
+    int T_inc = 2;
+    int T_lower = 2;
+    int T_upper = 200;
+    int alpha = 10;
+    if(argc == 11) {
+        N = stoi(argv[1]);
+        K = stoi(argv[2]);
+        R_incdec = stof(argv[3]);
+        R_lower = stoi(argv[4]);
+        R_scale = stoi(argv[5]);
+        T_dec = stof(argv[6]);
+        T_inc = stoi(argv[7]);
+        T_lower = stoi(argv[8]);
+        T_upper = stoi(argv[9]);
+        alpha = stoi(argv[10]);
+    }
+    VideoCapture cap1("dataset/Jackson_Hole_Wyoming/out0.mov");
+    VideoCapture cap2("dataset/Jackson_Hole_Wyoming/out3.mov");
+    VideoCapture cap3("videos/railway_s.mp4");
+
+    PBAS *pbas1 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
+    PBAS *pbas2 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
+    PBAS *pbas3 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
+    Mat frame1, frame2, frame3;
+    Mat *mask1, *mask2, *mask3;
+    
+    // process
+    while(1) {
+        cout << current_frame / FPS << "s" << endl;
+        cap1 >> frame1;
+        cap2 >> frame2;
+        cap3 >> frame3;
+        if (frame1.empty() || frame2.empty() || frame3.empty()) break;
+
+        GaussianBlur(frame1, frame1, Size(3,3), 0,0);
+        GaussianBlur(frame2, frame2, Size(3,3), 0,0);
+        GaussianBlur(frame3, frame3, Size(3,3), 0,0);
+
[…]

### 2019-02-08 pbas.cpp

@@ -81,7 +81,7 @@ float PBAS::distance(uint8_t a, uint8_t b) {
     return abs(a-b);
 }
 
-float PBAS::distance(uint8_t p, uint8_t p_grad, uint8_t g, uint8_t g_grad) {
+double PBAS::distance(uint8_t p, double p_grad, uint8_t g, double g_grad) {
     return (this->alpha/this->I_m) * abs((int)p_grad - (int)g_grad) + abs((int)p - (int)g); 
 }
 
@@ -152,8 +152,8 @@ Mat* PBAS::process(const Mat* frame) {
     this->frame_grad = gradient_magnitude(&this->frame);
     this->I_m = mean(this->frame_grad).val[0];
 
-    showCVMat(this->frame_grad, true, "grad");
-    moveWindow("grad", 900, 900);
+    // showCVMat(this->frame_grad, true, "grad");
+    // moveWindow("grad", 900, 900);
 
     // B, D, d_minavg initialization
     if (B.size() == 0) {
@@ -195,7 +195,7 @@ Mat* PBAS::process(const Mat* frame) {
     auto start = high_resolution_clock::now();
     for(int x=0; x < nRows; ++x) {
         this->i = this->frame.ptr<uint8_t>(x);
-        this->i_grad = frame_grad.ptr<uint8_t>(x);
+        this->i_grad = frame_grad.ptr<double>(x);
         this->q = F.ptr<uint8_t>(x);
         this->q_shadow_hsv = F_shadow_hsv.ptr<uint8_t>(x);
         this->r = R.ptr<float>(x);
@@ -220,7 +220,7 @@ Mat* PBAS::process(const Mat* frame) {
     auto duration = duration_cast<milliseconds>(stop - start);
     
     medianBlur(F,F,3);
-    cout << duration.count() << "ms" << endl;
+    //cout << duration.count() << "ms" << endl;
 
     // color_normalized_cross_correlation();
     //medianBlur(mask_shadows, mask_shadows, 3);
@@ -267,11 +267,11 @@ Mat* PBAS::process(const Mat* frame) {
 
     // showCVMat(T, true, "T");
 
-    showCVMat(R, true, "R");
-    moveWindow("R", 20, 340);
+    // showCVMat(R, true, "R");
+    // moveWindow("R", 20, 340);
 
[…]

### 2019-02-08 pbas.h

@@ -43,7 +43,7 @@ class PBAS
 
         const uint8_t *i;
         const Vec3b *i_rgb;
-        const uint8_t *i_grad;
+        const double *i_grad;
         uint8_t *q;
         uint8_t *q_shadow_hsv;
         float *r;
@@ -53,7 +53,7 @@ class PBAS
         void init();
 
         float distance(uint8_t, uint8_t);
-        float distance(uint8_t p, uint8_t p_grad, uint8_t g, uint8_t g_grad);
+        double distance(uint8_t p, double p_grad, uint8_t g, double g_grad);
 
         void updateMedian(int col);
         void updateF(int x, int y, int i_ptr);

### 2019-02-08 readvideo.cpp

@@ -122,7 +122,7 @@ int main(int argc, char const *argv[]) {
     Mat frame;
     Mat* mask;
     Mat edges;
-    float fps = cap.get(CV_CAP_PROP_FPS);
+    float fps = cap.get(5);
     float act_seconds = 0.0, partial_seconds = -1 / fps;
     int count_time_slot = 0;
     string folder_base, act_subfolder_base;

### 2019-02-10 comparevideos.cpp
PBAS rgb with bugs

@@ -2,7 +2,7 @@
 #include <string>
 #include <opencv2/opencv.hpp>
 #include <sys/stat.h> 
-#include "pbas.cpp"
+#include "pbas_rbg.cpp"
 //#include <limits>
 
 using namespace std;
@@ -42,12 +42,13 @@ int main(int argc, char const *argv[]) {
     PBAS *pbas1 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
     PBAS *pbas2 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
     PBAS *pbas3 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
+    pbas1->verbose = false; pbas2->verbose = false; pbas3->verbose = false;
     Mat frame1, frame2, frame3;
-    Mat *mask1, *mask2, *mask3;
+    Mat mask1, mask2, mask3;
     
     // process
     while(1) {
-        cout << current_frame / FPS << "s" << endl;
+        //cout << current_frame / FPS << "s" << endl;
         cap1 >> frame1;
         cap2 >> frame2;
         cap3 >> frame3;
@@ -57,24 +58,24 @@ int main(int argc, char const *argv[]) {
         GaussianBlur(frame2, frame2, Size(3,3), 0,0);
         GaussianBlur(frame3, frame3, Size(3,3), 0,0);
 
-        mask1 = pbas1->process(&frame1);
-        mask2 = pbas2->process(&frame2);
-        mask3 = pbas3->process(&frame3);
+        mask1 = pbas1->process(frame1);
+        mask2 = pbas2->process(frame2);
+        mask3 = pbas3->process(frame3);
         
         imshow("Jackson1", frame1);
-        moveWindow("Jackson1", 10,20);
-        imshow("Jackson1 mask", *mask1);
-        moveWindow("Jackson1 mask", 410, 20);
+        if(current_frame == 0) moveWindow("Jackson1", 10,20);
+        imshow("Jackson1 mask", mask1);
+        if(current_frame == 0) moveWindow("Jackson1 mask", 410, 20);
 
         imshow("Jackson2", frame2);
-        moveWindow("Jackson2", 10,300);
-        imshow("Jackson2 mask", *mask2);
-        moveWindow("Jackson2 mask", 410, 300);
[…]

### 2019-02-10 pbas.cpp

@@ -168,7 +168,7 @@ Mat* PBAS::process(const Mat* frame) {
             B_grad.push_back(b_grad_elem);
             //B_grad.push_back(frame_grad.clone());
 
-            Mat d_elem = Mat::zeros(h, w, CV_8UC1);
+            Mat d_elem = Mat::zeros(h, w, CV_32FC1);
             D.push_back(d_elem);
         }
         F = Mat::zeros(h, w, CV_8UC1);
@@ -370,17 +370,17 @@ void PBAS::updateB(int x, int y, int i_ptr) {
 void PBAS::updateR(int x, int y, int n, int i_ptr) {
     // find dmin
     uint8_t I = i[i_ptr];
-    uint8_t I_grad = i_grad[i_ptr];
-    int d_min = 255;
-    int d_act = 0;
+    float I_grad = i_grad[i_ptr];
+    float d_min = 255;
+    float d_act = 0;
     for (int i=0; i<N; i++){
-        d_act = distance(I, I_grad, B[i].at<uint8_t>(x, y), B_grad[i].at<uint8_t>(x, y));
+        d_act = distance(I, I_grad, B[i].at<uint8_t>(x, y), B_grad[i].at<float>(x, y));
         if (d_act < d_min)
             d_min = d_act;
     }
 
     // update Dk
-    D[n].at<uint8_t>(x, y) = d_min;
+    D[n].at<float>(x, y) = d_min;
 
     // find davg
     float d_cum = 0;

### 2019-02-10 pbas_rbg.cpp

@@ -1,0 +1,585 @@
+#include <iostream>
+#include <math.h>
+#include <stdlib.h>
+#include <time.h>
+#include <utility>      
+#include <algorithm>
+#include <opencv2/opencv.hpp>
+
+#include <chrono>
+using namespace std::chrono;
+
+using namespace std;
+using namespace cv;
+
+class PBAS
+{
+    private:
+        int frame_count = 0;
+
+        // model params
+        int N;
+        int K;
+        float R_incdec;
+        float R_lower;
+        float R_scale;
+        float T_dec;
+        float T_inc;
+        float T_lower;
+        float T_upper;
+        int alpha;
+        float I_m;
+        
+        // frame
+        Mat frame;
+        vector<Mat> frame_grad;
+        int w;
+        int h;
+
+        // model support matrices and pointers
+        const Vec3b *i;
+        const float *i_grad;
+
+        vector<vector<Mat>> B;
+        vector<vector<Mat>> B_grad;
+
+        vector<Mat> R; float *r;
+        vector<Mat> T; float *t;
+    
+        vector<vector<Mat>> D;
+        vector<Mat> d_minavg;
+
+        vector<pair<int,int>> displacement_vec;
+        
+        //Vec3b *med;
+
+        // output
+        vector<Mat> F; uint8_t *q;
+        //Mat final_mask;
+
+        // methods
+        void init(int channels);
+        double distance(uint8_t p, double p_grad, uint8_t g, double g_grad);
+        //void updateMedian(int col);
+
+        void updateF(int x, int y, int c);
+        void updateB(int x, int y, int c);
+        void updateR(int x, int y, int c, int n);
+        void updateR_notoptimized(int x, int y, int c, int n);
+        void updateT(int x, int y, int c);
+
+        void init_Mat(Mat &matrix, float initial_value);
+        Mat gradient_magnitude(const Mat &frame);
+
+    public:
+        PBAS();
+        PBAS(int N, int K, float R_incdec, int R_lower, int R_scale, float T_dec, int T_inc, int T_lower, int T_upper, int alpha);
+        ~PBAS();
+        
+        bool verbose = true;
+
[…]

### 2019-02-10 pbas_rbg.cpp
working on 3 channel distance abs

@@ -27,7 +27,7 @@ class PBAS
         float T_inc;
         float T_lower;
         float T_upper;
-        int alpha;
+        float alpha;
         float I_m;
         
         // frame
@@ -61,6 +61,7 @@ class PBAS
         void init(int channels);
         double distance(uint8_t p, double p_grad, uint8_t g, double g_grad);
         //void updateMedian(int col);
+        float distance_abs(uint8_t a, uint8_t b);
 
         void updateF(int x, int y, int c);
         void updateB(int x, int y, int c);
@@ -162,6 +163,10 @@ Mat PBAS::process(const Mat &frame) {
     //     return &F;
     // }
 
+    showCVMat(F[0], false, "f0");
+    showCVMat(F[1], false, "f1");
+    showCVMat(F[2], false, "f3");
+
     auto start = high_resolution_clock::now();
     for(int c=0; c < channels; c++) {
         for(int x=0; x < this->h; x++) {
@@ -175,16 +180,18 @@ Mat PBAS::process(const Mat &frame) {
             this->I_m = mean(this->frame_grad[c]).val[0];
 
             for (int y=0; y < this->w; y++) {
+                //cout << "x,y" << x << ";" << y << endl;
                 updateF(x, y, c);
-                // updateT(x, y, c);
+                updateT(x, y, c);
                 // q[y] = i[y][0];
             }
+            
         }
         // showCVMat(frame_grad[c], false, format("framegrad%d",c));
         // showCVMat(R[c], false, format("R%d",c));
-        showCVMat(F[c], false, format("F%d",c));
+        //showCVMat(F[c], false, format("F%d",c));
         // showCVMat(T[c], false, format("T%d",c));
-        medianBlur(F[c],F[c],3);
+        //medianBlur(F[c],F[c],3);
     }
 
     if(verbose) {
@@ -209,7 +216,7 @@ void PBAS::init(int channels) {
             Mat b_elem(h, w, CV_8UC1);
             randu(b_elem, 0, 255);
             b.push_back(b_elem);
-            //b.push_back(this->frame.clone());
+            //b.push_back(this->frame[c].clone());
 
             Mat b_grad_elem(h, w, CV_32FC1);
[…]

### 2019-02-10 pbas_rbg.cpp
Vector pointers to support parallel access

@@ -28,23 +28,23 @@ class PBAS
         float T_lower;
         float T_upper;
         int alpha;
-        float I_m;
+        vector<float> I_m;
         
         // frame
-        Mat frame;
+        vector<Mat> frame;
         vector<Mat> frame_grad;
         int w;
         int h;
 
         // model support matrices and pointers
-        const Vec3b *i;
-        const float *i_grad;
+        vector<uint8_t*> i;
+        vector<float*> i_grad;
 
         vector<vector<Mat>> B;
         vector<vector<Mat>> B_grad;
 
-        vector<Mat> R; float *r;
-        vector<Mat> T; float *t;
+        vector<Mat> R; vector<float*> r;
+        vector<Mat> T; vector<float*> t;
     
         vector<vector<Mat>> D;
         vector<Mat> d_minavg;
@@ -54,12 +54,12 @@ class PBAS
         //Vec3b *med;
 
         // output
-        vector<Mat> F; uint8_t *q;
+        vector<Mat> F; vector<uint8_t*> q;
         //Mat final_mask;
 
         // methods
         void init(int channels);
-        double distance(uint8_t p, double p_grad, uint8_t g, double g_grad);
+        double distance(double p, double p_grad, double g, double g_grad, int c);
         //void updateMedian(int col);
 
         void updateF(int x, int y, int c);
@@ -100,7 +100,7 @@ PBAS::PBAS(int N, int K=2, float R_incdec=0.05, int R_lower=18, int R_scale=5, f
     this->T_lower = T_lower;
     this->T_upper = T_upper;
     this->alpha = alpha;
-    this->I_m = 1.0;
+    //this->I_m = 1.0;
     
     //initialize random seed
     //srand (time(NULL));
@@ -117,7 +117,7 @@ PBAS::PBAS(int N, int K=2, float R_incdec=0.05, int R_lower=18, int R_scale=5, f
 }
 
 PBAS::~PBAS() {
-    frame.release();
+    frame.clear();
     frame_grad.clear();
     //median.release();
     B.clear();
@@ -139,7 +139,8 @@ void PBAS::init_Mat(Mat &matrix, float initial_value){
 }
 
 Mat PBAS::process(const Mat &frame) {
-    this->frame = frame;
[…]

### 2019-02-10 comparevideos.cpp
PBAS rbg semi-working (problems with gradients)

@@ -42,7 +42,7 @@ int main(int argc, char const *argv[]) {
     PBAS *pbas1 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
     //PBAS *pbas2 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
     //PBAS *pbas3 = new PBAS(N, K, R_incdec, R_lower, R_scale, T_dec, T_inc, T_lower, T_upper, alpha);
-    //pbas1->verbose = false; pbas2->verbose = false; pbas3->verbose = false;
+    pbas1->verbose = false; // pbas2->verbose = false; pbas3->verbose = false;
     Mat frame1, frame2, frame3;
     Mat mask1, mask2, mask3;
     
@@ -62,11 +62,11 @@ int main(int argc, char const *argv[]) {
         //mask2 = pbas2->process(frame2);
         //mask3 = pbas3->process(frame3);
         
-/*         imshow("Jackson1", frame1);
-        if(current_frame == 0) moveWindow("Jackson1", 10,20);
+        imshow("Jackson1", frame1);
+        if(current_frame == 0) moveWindow("Jackson1", 0,20);
         imshow("Jackson1 mask", mask1);
-        if(current_frame == 0) moveWindow("Jackson1 mask", 410, 20);
-
+        if(current_frame == 0) moveWindow("Jackson1 mask", 420, 20);
+/*
         imshow("Jackson2", frame2);
         if(current_frame == 0) moveWindow("Jackson2", 10,300);
         imshow("Jackson2 mask", mask2);
@@ -76,8 +76,6 @@ int main(int argc, char const *argv[]) {
         if(current_frame == 0) moveWindow("Railway", 10,580);
         imshow("Railway mask", mask3);
         if(current_frame == 0) moveWindow("Railway mask", 410, 580); */
-
-        imshow("final mask", mask1);
 
         char c=(char)waitKey(25);
         if(c==27) break;

### 2019-02-10 pbas_rbg.cpp

@@ -163,10 +163,6 @@ Mat PBAS::process(const Mat &frame) {
     //     return &F;
     // }
 
-    showCVMat(F[0], false, "f0");
-    showCVMat(F[1], false, "f1");
-    showCVMat(F[2], false, "f3");
-
     auto start = high_resolution_clock::now();
     for(int c=0; c < channels; c++) {
         for(int x=0; x < this->h; x++) {
@@ -246,7 +242,6 @@ void PBAS::init(int channels) {
         t.push_back(0);
         I_m.push_back(1);
     }
-    //this->frame_grad.resize(channels);
     assert(B.size() == channels);
     assert(B_grad.size() == channels);
     assert(D.size() == channels);
@@ -281,7 +276,8 @@ Mat PBAS::gradient_magnitude(const Mat &frame){
 }
 
 double PBAS::distance(double p, double p_grad, double g, double g_grad, int c) {
-    return (this->alpha/this->I_m[c]) * abs(p_grad - g_grad) + abs(p - g);
+    // TO-DO
+    // return (this->alpha/this->I_m[c]) * abs(p_grad - g_grad) + abs(p - g);
     //cout << abs(p - g) << endl;
     return abs(p - g);
 }
@@ -290,16 +286,15 @@ void PBAS::updateF(int x, int y, int c) {
     int k = 0;  // number of lower-than-R distances found so far
     int j = 0;
     while(j < N && k < K) {
+        // TO-DO
         if(distance(i[c][y], i_grad[c][y], B[c][j].at<uint8_t>(x,y), B_grad[c][j].at<float>(x,y), c) < r[c][y]) {
             k++;
         }
         j++;
     }
     // check if at least K distances are less than R(x,y) => background pixel
-    //cout << k << endl;
     if(k >= K) {
         q[c][y] = 0;
-        //this->F[c].at<uint8_t>(x,y)=0;
         // q_shadow_hsv[i_ptr]=0;
         updateB(x, y, c);
     } else {

### 2019-02-11 pbas_rbg.cpp
Multithreading

@@ -5,6 +5,7 @@
 #include <utility>      
 #include <algorithm>
 #include <opencv2/opencv.hpp>
+#include <thread>
 
 #include <chrono>
 using namespace std::chrono;
@@ -16,6 +17,8 @@ class PBAS
 {
     private:
         int frame_count = 0;
+
+        vector<thread> threads;
 
         // model params
         int N;
@@ -62,6 +65,8 @@ class PBAS
         double distance(double p, double p_grad, double g, double g_grad, int c);
         //void updateMedian(int col);
 
+
+        void process_channel(int c);
         void updateF(int x, int y, int c);
         void updateB(int x, int y, int c);
         void updateR(int x, int y, int c, int n);
@@ -127,6 +132,9 @@ PBAS::~PBAS() {
     F.clear();
     d_minavg.clear();
     displacement_vec.clear();
+    
+    for(thread &t : threads) t.join();
+    threads.clear();
 }
 
 
@@ -164,31 +172,12 @@ Mat PBAS::process(const Mat &frame) {
     // }
 
     auto start = high_resolution_clock::now();
+    threads.clear();
     for(int c=0; c < channels; c++) {
-        for(int x=0; x < this->h; x++) {
-            this->i[c] = this->frame[c].ptr<uint8_t>(x);
-            this->i_grad[c] = frame_grad[c].ptr<float>(x);
-            this->q[c] = F[c].ptr<uint8_t>(x);
-            this->r[c] = R[c].ptr<float>(x);
-            this->t[c] = T[c].ptr<float>(x);
-            //this->med = median.ptr<Vec3b>(x);
-
-            this->I_m[c] = mean(this->frame_grad[c]).val[0];
-
-            for (int y=0; y < this->w; y++) {
-                //cout << "x,y" << x << ";" << y << endl;
-                updateF(x, y, c);
-                updateT(x, y, c);
-                // q[y] = i[y][0];
-            }
-            
-        }
-        // showCVMat(frame_grad[c], false, format("framegrad%d",c));
-        // showCVMat(R[c], false, format("R%d",c));
-        // showCVMat(F[c], false, format("F%d",c));
-        // showCVMat(T[c], false, format("T%d",c));
-        //medianBlur(F[c],F[c],3);
+        threads.push_back(thread(&PBAS::process_channel, this, c));
     }
+    // wait all threads
[…]

### 2019-02-11 pbas_rbg.cpp
fixed distances with gradients

@@ -38,7 +38,7 @@ class PBAS
 
         // model support matrices and pointers
         vector<uint8_t*> i;
-        vector<float*> i_grad;
+        vector<uint8_t*> i_grad;
 
         vector<vector<Mat>> B;
         vector<vector<Mat>> B_grad;
@@ -167,7 +167,7 @@ Mat PBAS::process(const Mat &frame) {
     for(int c=0; c < channels; c++) {
         for(int x=0; x < this->h; x++) {
             this->i[c] = this->frame[c].ptr<uint8_t>(x);
-            this->i_grad[c] = frame_grad[c].ptr<float>(x);
+            this->i_grad[c] = frame_grad[c].ptr<uint8_t>(x);
             this->q[c] = F[c].ptr<uint8_t>(x);
             this->r[c] = R[c].ptr<float>(x);
             this->t[c] = T[c].ptr<float>(x);
@@ -276,10 +276,8 @@ Mat PBAS::gradient_magnitude(const Mat &frame){
 }
 
 double PBAS::distance(double p, double p_grad, double g, double g_grad, int c) {
-    // TO-DO
-    // return (this->alpha/this->I_m[c]) * abs(p_grad - g_grad) + abs(p - g);
-    //cout << abs(p - g) << endl;
-    return abs(p - g);
+    return (this->alpha/this->I_m[c]) * abs(p_grad - g_grad) + abs(p - g);
+    // return abs(p - g);
 }
 
 void PBAS::updateF(int x, int y, int c) {
@@ -287,7 +285,7 @@ void PBAS::updateF(int x, int y, int c) {
     int j = 0;
     while(j < N && k < K) {
         // TO-DO
-        if(distance(i[c][y], i_grad[c][y], B[c][j].at<uint8_t>(x,y), B_grad[c][j].at<float>(x,y), c) < r[c][y]) {
+        if(distance(i[c][y], i_grad[c][y], B[c][j].at<uint8_t>(x,y), B_grad[c][j].at<uint8_t>(x,y), c) < r[c][y]) {
             k++;
         }
         j++;
@@ -322,7 +320,6 @@ void PBAS::updateB(int x, int y, int c) {
         B_grad[c][n].at<float>(x, y) = i_grad[c][y];
     }
     
-    return;
     //generate a random number between 0 and 99
     rand_numb = rand() %100;
     if(rand_numb < update_p) {

### 2019-02-12 paper/Paper.tex
typos fixing

@@
 consists in finding a category for every pixel in a single or (like in this case) in a frames sequence
-of a video and saying if that pixels are part of background or foreground. At the end, the algorithm take as
-as input a video and output a binary mask, where 1s are foreground pixels and 0 background pixels.
+of a video and saying if that pixels are part of background or foreground.
+At the end, the algorithm take as input a video and output a binary mask, where 1s are foreground pixels and 0 background pixels.
 We restricted our domain to traffic monitoring, in particular this work should put the basis for an automated
@@
 It can be categorized as a non-parametric method, since it uses a history of N image values as the background
-model, and uses a random update rule similar to the one used by the ViBe algorithm. However, in Vibe, the
+model, and uses a random update rule similar to the one used by the ViBe algorithm. However, in ViBe, the
 randomness parameters as well as the decision threshold are fixed for all pixels. In contrast, in the PBAS
@@
 This section describes the Pixel-Based Adaptive Segmenter, which follows a non-parametric paradigm. Thus,
-every pixel x i is modeled by an array of recently observed background values.
+every pixel $x_i$ is modeled by an array of recently observed background values.
 The method consists of several components which are depicted as a state machine in Figure 1.

### 2019-02-12 paper/Paper.tex
paragraphs + distance in pbas description

@@
 the foreground segmentation mask is calculated as:
-\begin{equation}
+\begin{equation} \label{eq:F}
 F(x_i) = \begin{cases} 1,& \#\{dist(I(x_i), B_k(x_i)) < R(x_i)\} < K \\0 &else\end{cases}
@@
 $F = 1$ implies foreground.
-The decision involves two parameters: (1) The distance threshold $R(x_i)$,which is defined for each pixel separately and which can change dynamically; and (2) the minimum number $K$, which is a fixed global parameter.\newline
+The decision involves two parameters: the distance threshold $R(x_i)$,which is defined for each pixel separately and which can change dynamically; and the minimum number $K$, which is a fixed global parameter.
+\paragraph{Update B}
 The background model is only updated for those pixels that
 are currently background (i.e. $F(x_i) = 0$).
-Updating means that for a certain index $k \in \{1...N \}$(chosen uniformly at random), the corresponding background model
+Updating means that for a certain index $k \in \{1...N \}$ (chosen uniformly at random), the corresponding background model
 value $B_k(x_i)$ is replaced by the current pixel value $I(x_i)$.
 This allows the current pixel value to be ”learned” into
-the background model. This update, however, is only performed with probability $p = 1/T(x_i)$.Otherwise no update
+the background model. This update, however, is only performed with probability $p = 1/T(x_i)$. Otherwise no update
 is carried out at all. Therefore, the parameter $T(x_i)$ defines
@@
 will be updated.
-Is also update (with probability $p = 1/T(x_i)$) a randomly chosen neighboring pixel $y i \in N(x_i)$. Thus, the
-background model $B_k(y_i)$ at this neighboring pixel is replaced by its current pixel value $V(y_i)$. \newline
+It is also updated (with probability $p = 1/T(x_i)$) a randomly chosen neighboring pixel $y_i \in N(x_i)$. Thus, the
+background model $B_k(y_i)$ at this neighboring pixel is replaced by its current pixel value $V(y_i)$.
+\paragraph{Update R}
 To change the decision treshold $R(x_i)$ is saved an array of recently observed pixel values in the background model $B(x_i)$, we also create an array $\mathbf{D}(x_i) = \{D_1(x_i ),...,D_N(x_i)\}$ of minimal decision distances.
-Whenever an update of $B_k(x_i)$ is carried out, the currently observed minimal distance $d_{min}(x_i) = min_k dist(I(x_i), B_k(x_i))$ is written to this array: $D_k(x_i) \leftarrow d_{min}(x_i)$.
+Whenever an update of $B_k(x_i)$ is carried out, the currently observed minimal distance:
+\[ d_{min}(x_i) = min_k dist(I(x_i), B_k(x_i)) \] is written to this array: $D_k(x_i) \leftarrow d_{min}(x_i)$.
 Thus, a history of minimal decision distances is created.
@@
 Here, $R_{inc/dec}$, $R_{scale}$ are fixed parameters. This can be seen as a dynamic controller for the state variable $R(x_i)$.
-\newline
+\paragraph{Update T}
 Independent of the foreground state $F(x_i)$, every object will be merged into the background depending on the learning parameter
@@
 $T_{inc}$ and $T_{dec}$ are fixed parameters. Are also defined $T_{lower}$ and $T_{upper}$ such that the values can not go out of a specific bound.
+\paragraph{Distance computation}
+The quantity $ dist(I(x_i),B_k(x_i)) $ is computed as:
+\begin{equation}
+\frac{\alpha}{\overline{I^m}}|I^m(x_i) - B_k^m(x_i)| + |I^v(x_i) - B_k^v(x_i)|
+\end{equation}
+where $m$ indicates the gradient magnitude and $v$ the pixel intensity value.
+$\overline{I_m}$ is the average gradient magnitude over the last observed frame.
+Thus, the fraction $\frac{\alpha}{\overline{I^m}}$ weighs the importance
+of pixel values against the gradient magnitude
 

### 2019-02-15 paper/Paper.tex
Added section T in the paper

@@
 
-\subsection*{T max}
-$T_{max}$ controls the lowest probability of updating a pixel and putting it the
-background model $B$.
+\subsection*{T}
+The controller $T(x_i)$ associated to each pixel, is introduced with the idea to change dinamically the learning rate associated to each pixel based on its classification.
+It is slowly increased when the pixel is classified as background and slowly decreased when the pixel is classified as foreground.
+That leads to have a better foreground detection for multiple reasons.
+\newline
+In case of highly dynamic background (i.e.
+big $\overline{d}_{min}(x_i)$), the learning parameter $T(x_i)$ stays constant or only slightly changes.
+In this case of highly dynamic background, erroneously detected foreground will not remain for long, be-
+cause the probability to update $p = 1/T(x_i)$ does not reach
+zero too fast. \newline
+In the other ideal case of a fully static background, a classification as foreground is quite solid. \newline
+$T_{upper}, T_{lower}:$ control respectively the lowest and the highest probability of updating a pixel and putting it in the
+background model $B$. \newline
+$T_{inc}, T_{dec}:$ are fixed parameters and control the update of the controller T of each pixel, has been chosen to differentiate the two quantity since the pixels that are background are usually the largest part so choosing different parameters can lead to a better perfomance.\newline
+Since one of the objective of the work is to have a robust classification of the foreground objects, we want to find the best values for the T parameters such that an object if classified as foreground stay as it for the longest period of time.
+INSERIRE IMMAGINI
 

### 2019-02-15 pbas_rbg.cpp
Bg estimation

@@ -56,15 +56,13 @@ class PBAS
         
         //Vec3b *med;
 
-        // output
+        // mask
         vector<Mat> F; vector<uint8_t*> q;
-        //Mat final_mask;
 
         // methods
-        void init(int channels);
+        void init(const Mat &frame);
         double distance(double p, double p_grad, double g, double g_grad, int c);
         //void updateMedian(int col);
-
 
         void process_channel(int c);
         void updateF(int x, int y, int c);
@@ -90,6 +88,7 @@ class PBAS
         void showCVMat(Mat matrix, bool normalize, string window_name);
 
         //Mat median;
+        Mat estimatedBg;
 };
 
 
@@ -125,6 +124,7 @@ PBAS::~PBAS() {
     frame.clear();
     frame_grad.clear();
     //median.release();
+    estimatedBg.release();
     B.clear();
     R.clear();
     D.clear();
@@ -148,14 +148,13 @@ void PBAS::init_Mat(Mat &matrix, float initial_value){
 
 Mat PBAS::process(const Mat &frame) {
     split(frame, this->frame);
-    //this->frame = frame;
     this->w = frame.cols;
     this->h = frame.rows;
     int channels = frame.channels();
 
     // B, D, d_minavg initialization
     if (B.size() == 0) {
-        init(channels);
+        init(frame);
     }
     
     // gradients computation
@@ -178,7 +177,6 @@ Mat PBAS::process(const Mat &frame) {
     }
     // wait all threads
     for(thread &t : threads) t.join();
-    //showCVMat(B[0][0], false, format("B%d",0));
 
     if(verbose) {
         auto stop = high_resolution_clock::now();
@@ -186,13 +184,14 @@ Mat PBAS::process(const Mat &frame) {
         cout << duration.count() << "ms" << endl;
     }
 
-    Mat res = Mat::zeros(h, w, CV_8UC1);
+    Mat final_mask = Mat::zeros(h, w, CV_8UC1);
     for(int c = 0; c < channels; c++) {
-        bitwise_or(res, F[c], res);
+        bitwise_or(final_mask, F[c], final_mask);
     }
+    frame.copyTo(estimatedBg, 255-final_mask);
     // medianBlur(res,res,3);
     
-    return res;
+    return final_mask;
 }
 
 void PBAS::process_channel(int c) {
@@ -220,8 +219,8 @@ void PBAS::process_channel(int c) {
 }
 
[…]

### 2019-02-15 pbas_rbg.cpp
Optical flow bg estimation

@@ -32,12 +32,16 @@ class PBAS
         float T_upper;
         float alpha;
         vector<float> I_m;
+
+        float flow_mag_threshold = 0.7;
+        int consecutive_flow_activation = 6;
         
         // frame
         vector<Mat> frame;
         vector<Mat> frame_grad;
         int w;
         int h;
+        Mat previous_frame;
 
         // model support matrices and pointers
         vector<uint8_t*> i;
@@ -48,6 +52,10 @@ class PBAS
 
         vector<Mat> R; vector<float*> r;
         vector<Mat> T; vector<float*> t;
+
+        // flow
+        //Mat flow_active;
+        vector<Mat> flow_active_history;
     
         vector<vector<Mat>> D;
         vector<Mat> d_minavg;
@@ -60,7 +68,7 @@ class PBAS
         vector<Mat> F; vector<uint8_t*> q;
 
         // methods
-        void init(const Mat &frame);
+        void init(int channels, int frametype);
         double distance(double p, double p_grad, double g, double g_grad, int c);
         //void updateMedian(int col);
 
@@ -71,6 +79,8 @@ class PBAS
         void updateR_notoptimized(int x, int y, int c, int n);
         void updateT(int x, int y, int c);
 
+        void analyzeOpticalFlow(const Mat &frame);
+
         void init_Mat(Mat &matrix, float initial_value);
         Mat gradient_magnitude(const Mat &frame);
 
@@ -80,6 +90,7 @@ class PBAS
         ~PBAS();
         
         bool verbose = true;
+        bool use_optical_flow = true;
 
         // main
         Mat process(const Mat &frame);
@@ -89,6 +100,7 @@ class PBAS
 
         //Mat median;
         Mat estimatedBg;
+        Mat flow_active;
 };
 
 
@@ -154,7 +166,9 @@ Mat PBAS::process(const Mat &frame) {
 
     // B, D, d_minavg initialization
     if (B.size() == 0) {
-        init(frame);
+        init(channels, frame.type());
+        
+        cvtColor(frame, previous_frame, COLOR_BGR2GRAY);
     }
     
     // gradients computation
@@ -175,6 +189,7 @@ Mat PBAS::process(const Mat &frame) {
     for(int c=0; c < channels; c++) {
[…]

### 2019-02-16 paper/Paper.tex
Typo
black refs

--- 2019-02-12 paper/Paper.tex
@@
 the decision threshold can be dynamically adapted as follows:
-\begin{equation}
+\begin{equation} \label{eq:R}
 R(x_i)=\begin{cases} R(x_i)\cdot(1-R{inc/dec}) &if\ R(x_i)> \overline{d}_{min}(x_i)\cdot R_{scale} \\ R(x_i)\cdot (1+R{inc/dec}) &else  \end{cases}
@@
 The quantity $ dist(I(x_i),B_k(x_i)) $ is computed as:
-\begin{equation}
+\begin{equation} \label{eq:dist}
 \frac{\alpha}{\overline{I^m}}|I^m(x_i) - B_k^m(x_i)| + |I^v(x_i) - B_k^v(x_i)|
@@
 \subsection*{R}
-Eq ?????????? is similar to the equation of a pure proportional controller.
-In this case the $R(x,y)$ has to follow $d_{minavg}(x,y) * R_{scale}$. So:
+Eq \ref{eq:R} is similar to the equation of a pure proportional controller.
+In this case the $R(x,y)$ has to follow $d_{minavg}(x,y)R_{scale}$. So:
 \begin{itemize}
@@
 \subsection*{alpha}
-From eq ?????????? we see that alpha tells how much to weight the image gradients
+From eq \ref{eq:dist} we see that alpha tells how much to weight the image gradients
 with respect to the image values in the distance computation.

### 2019-02-16 pbas.h
gradients read fine in pbas greyscale

@@ -43,7 +43,7 @@ class PBAS
 
         const uint8_t *i;
         const Vec3b *i_rgb;
-        const double *i_grad;
+        const uint8_t *i_grad;
         uint8_t *q;
         uint8_t *q_shadow_hsv;
         float *r;
@@ -53,7 +53,7 @@ class PBAS
         void init();
 
         float distance(uint8_t, uint8_t);
-        double distance(uint8_t p, double p_grad, uint8_t g, double g_grad);
+        float distance(uint8_t p, uint8_t p_grad, uint8_t g, uint8_t g_grad);
 
         void updateMedian(int col);
         void updateF(int x, int y, int i_ptr);

### 2019-02-16 pbas_rbg.cpp
Optical flow

@@ -55,6 +55,7 @@ class PBAS
 
         // flow
         //Mat flow_active;
+        vector<uint8_t*> fw;
         vector<Mat> flow_active_history;
     
         vector<vector<Mat>> D;
@@ -203,8 +204,12 @@ Mat PBAS::process(const Mat &frame) {
     for(int c = 0; c < channels; c++) {
         bitwise_or(final_mask, F[c], final_mask);
     }
+    medianBlur(final_mask,final_mask,3);
+
     frame.copyTo(estimatedBg, 255-final_mask);
-    // medianBlur(res,res,3);
+
+    Mat toshow; T[0].convertTo(toshow, CV_8UC1);
+    imshow("T", toshow); moveWindow("T", 0, 550);
 
     return final_mask;
 }
@@ -217,8 +222,8 @@ void PBAS::process_channel(int c) {
         this->r[c] = R[c].ptr<float>(x);
         this->t[c] = T[c].ptr<float>(x);
         //this->med = median.ptr<Vec3b>(x);
-
         this->I_m[c] = mean(this->frame_grad[c]).val[0];
+        this->fw[c] = flow_active.ptr<uint8_t>(x);
 
         for (int y=0; y < this->w; y++) {
             updateF(x, y, c);
@@ -226,11 +231,6 @@ void PBAS::process_channel(int c) {
         }
         
     }
-    // showCVMat(frame_grad[c], false, format("framegrad%d",c));
-    // showCVMat(R[c], false, format("R%d",c));
-    // showCVMat(F[c], false, format("F%d",c));
-    // showCVMat(T[c], false, format("T%d",c));
-    //medianBlur(F[c],F[c],3);
 }
 
 // initialize all support matrices
@@ -269,6 +269,7 @@ void PBAS::init(int channels, int frametype) {
         r.push_back(0);
         t.push_back(0);
         I_m.push_back(1);
+        fw.push_back(0);
     }
     estimatedBg = Mat::ones(h, w, frametype);
 
@@ -428,6 +429,12 @@ void PBAS::updateT(int x, int y, int c) {
     else
         t[c][y] -= T_dec / (d_minavg[c].at<float>(x,y) + 0.1);
 
+    if(use_optical_flow) {
+        if(fw[c][y] == 0) t[c][y] -= 10;
+    }
+    
[…]

### 2019-02-16 paper/Paper.tex

--- 2019-02-15 paper/Paper.tex
@@
 Each of them stores the history of pixels values.
-The greater, the more complex background can be handled but at the cost of heavier
+The greater, the more complex backgrounds can be handled but at the cost of heavier
 computation. At $N=35$, experimental results show that performance saturates.

### 2019-02-16 paper/Paper.tex
Added parameter settings

@@
 
+\subsection{parameter settings}
+The proposed method has a multitude of tunable parameters, their values have been chosen focusing the attention on having an algorithm robust to changing on weather condition and capable of identifying foreground object even if they stay steady for a period of time (i.e.
+cars stopped to a traffic light).
+\newline
+Experimental results have been shown that the optimal parameters are the following:
+\begin{itemize}
+\item $N = 30:$ The number of components of the
+background model.
+Increasing the value of this parameter let the algorithm be more robust to noise, but at the same time, increase a lot the time complexity.
+The value has been choosen considering the previous tradeoff.
+\item $K = 3:$ The number of components that have to
+be closer than $R(x_i)$ in order to set the pixel to be background.
+\item $R_{incdec} = 0.05:$ The rate, at which the decision
+threshold $R(x_i)$ is regulated by the controller. 
+\item $R_{lower}=18:$ Lower bound of the decision threshold. It has been left to the default value.
+\item $R_{scale}=2$ Scaling factor in the controller for the
+decision threshold. Lowering it from 5 to 2 has shown to let the algorithm to recognize better the foreground object introducing very low noise.
+\item $T_{dec} = 0.05:$ If $x_i$ is background, $T_{dec}$ is the rate at which $T(x_i)$ is decreased.
+\item $T_{inc} = 5:$ If $x_i$ is foreground, $T_{inc}$ is the rate at which $T(x_i)$ is increased.
+These parameter has shown to be very important to keep classified as foreground pixels associated to objects that have stopped their movement.
+Increasing it to 5 from the proposed value of 2 reduces faster the probability to put a steady foreground object in to one of the N components of the background model.
+\item $T_{lower}= 2:$ Lower bound of $T(x_i)$.
+\item $T_{upper}= 200:$ Upper bound of $T(x_i)$.
+\end{itemize}
+
 \subsection*{Conclusion}

### 2019-02-16 paper/Paper.tex
paper conclusions

--- 2019-02-16 paper/Paper.tex
@@
 \subsection*{Conclusion}
+We have provided a background subtraction system based on the PBAS algorithm.
+The system is meant to work real time in traffic surveillance scenarios, also under challening environmental conditions and heavy traffic, 24/7.
+In order to do this, the general purpose PBAS algorithm was studied and the hyperparameters were adapted in order to perform best in this particular scenario.
+We have validated our work by visual inspection in particular and challenging situations, with long videos with environmental and traffic conditions changing over time.
+We further experimented with shadow removal methods and we tried to exploit motion informations.
+Future work can be done in this direction, until arriving to a precise classification of noise and errors in the mask detected.
+Still, the system as it is now can be employed as a preprocessing step for many anomaly detection applications, such as car accidents, cars going in the wrong direction and people crossing the road.

### 2019-02-17 paper/Paper.tex
various paper things + default parameters on the algorithm

@@
 At the end have been studied how to enhanche the performance of the algorithm solving the issue of the presence of the shadows in the video, using techniques of shadows removal.
-\subsubsection{PBAS algorithm}
+\subsection{PBAS algorithm} \label{subsect:PBAS}
 
@@
 \section*{Experiments}
-In this section we first report the results obtained by varying the main model parameters. We also include the
-adopted settings that perform best in our scenario. Then, we show some real-case usages of the algorithm,
-comparing the different environmental conditions that we found in 24/24h live videos.
+
+In this section we first report the results obtained by varying the main model parameters.
+We also include the adopted settings that perform best in our scenario.
+Then, we show some real-case usages of the algorithm, comparing different environmental conditions that we found in 24/24h live videos.
+Next, are reported experiments made with
+
+\subsection{Hyperparameters}
 
@@
 \begin{itemize}
-    \item $R_{incdec}$: setting too high lead to oscillation around the target and if
-    set too high leads to divergence. The default value is 0.05 which leads to fast
-    enough adaptation to the target value (in more or less 30 frames the target is
-    reached). Assuming that the backgroud isnt changing much there is no reason to
-    raise it.
+    \item $R_{incdec}$: it can be seen as the propositional constant of the controller: setting too high lead to oscillation around the target and if set too high leads to divergence.
+Setting too low leads to slow convergence to the target values, making the algorithm too slow to adapt to dynamic backgrounds.
     \item $R_{scale}$: tells how much the algorithm is sensitive, in fact it is a direct
@@
     }
+    \caption{As $R_{scale}$ grows the tree leaves that are waved by the wind are not detected
+anymore as foreground but also a part of the car is lost.}
 \end{figure}
-As we see, as $R_{scale}$ grows the tree leaves that are waved by the wind are not detected
-anymore as foreground but also a part of the car is lost.
 
-\subsection*{alpha}
-From eq \ref{eq:dist} we see that alpha tells how much to weight the image gradients
+\subsection*{$\alpha$}
+From eq \ref{eq:dist} we see that $\alpha$ tells how much to weight the image gradients
 with respect to the image values in the distance computation.
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.45\textwidth]{Figures/gradients.jpg}
-    \caption{Gradients}
-    \label{fig:gradients}
-\end{figure}
+%\begin{figure}[!t]
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{Figures/gradients.jpg}
+%    \caption{Gradients}
+%    \label{fig:gradients}
+%\end{figure}
 As we see, the image gradients give a good cue to distinguish the street from the cars
@@
     }
+    \caption{The influence of the image gradients.
+In \textbf{(a)} is reported the mask not considering the gradients ($\alpha = 0$).
+Going toward the bottom, we see the good influence on the detection, but we also witness an increase of the noise.}
 \end{figure}
 
-\subsection{parameter settings}
+\subsection{Optimal Setting}
 The proposed method has a multitude of tunable parameters, their values have been chosen focusing the attention on having an algorithm robust to changing on weather condition and capable of identifying foreground object even if they stay steady for a period of time (i.e.
-cars stopped to a traffic light).
+cars stopped to a traffic light), still guaranteeing a real time behaviour.
 \newline
-Experimental results have been shown that the optimal parameters are the following:
+Experimental results have shown that the optimal parameters are the following:
 \begin{itemize}
@@
 The value has been choosen considering the previous tradeoff.
-\item $K = 3:$ The number of components that have to
-be closer than $R(x_i)$ in order to set the pixel to be background.
-\item $R_{incdec} = 0.05:$ The rate, at which the decision
-threshold $R(x_i)$ is regulated by the controller. 
-\item $R_{lower}=18:$ Lower bound of the decision threshold. It has been left to the default value.
-\item $R_{scale}=2$ Scaling factor in the controller for the
-decision threshold. Lowering it from 5 to 2 has shown to let the algorithm to recognize better the foreground object introducing very low noise.
+\item $K = 3:$ The number of components that have to be closer than $R(x_i)$ in order to set the pixel to be background. It has been left to the value proposed in the original paper.
+\item $R_{incdec} = 0.05:$ Proportial constant of the controller for the variable $R(x_i)$.
+The value 0.05 leads to pretty fast adaptation to the target value of $R(x_i)$ (in more or less 30 frames the target is reached).
+Considering that our target are long periods of time there is no reason to raise it.
+\item $R_{lower}=18:$ Lower bound of the decision threshold.
+It has been left to the value proposed in the original paper, since we found that raising it would lead to a significant decrease in true positive and lowering it would let too much noise in.
+\item $R_{scale}=2$ Scaling factor in the controller for the decision threshold.
+Setting it to 2 has shown to let the algorithm to recognize better the foreground object while introducing very low noise.
 \item $T_{dec} = 0.05:$ If $x_i$ is background, $T_{dec}$ is the rate at which $T(x_i)$ is decreased.
@@
 \item $T_{upper}= 200:$ Upper bound of $T(x_i)$.
+\item $\alpha = 20:$ Weight to give to image gradients. The proposed value is high because this shows to be a good cue for distinguishing car objects
 \end{itemize}
+
+\subsection{RGB vs Grey}
+We have experimented with a version of the algorithm working just on grey scale images and with one exploiting also color informations.
+The grey one is implementing exactly all the steps described in \ref{subsect:PBAS}.
+The RGB version is running the Grey version for any of the three channels and then the final mask is computed as the Logical-OR of the three masks obtained for each of the three channels.
+Experiments show that the RGB version has a better cars detection capabilities, but it proves to be computationally heavy, breaking totally the real time capabilities that we manage to achieve.
+Moreover, the behaviour proves to be better, but it doesn't justify the cost of having a non-realtime algorithm.
+
+\begin{figure}[!t]
+    \centering
+    \subfloat[][Grey]{
+        \includegraphics[width=0.45\textwidth]{Figures/grey.png}
+    }
+    \newline
+    \subfloat[][RGB]{
+        \includegraphics[width=0.45\textwidth]{Figures/rgb.png}
+    }
+	\caption{RGB version compared with the Grey one. As we see, there is some difference present but it is negligible}
+\end{figure}
+
+\subsection{Long Period Experiments}
+
+\subsection{Algorithm Validation}
 

### 2019-02-17 readvideo.cpp

@@ -99,12 +99,12 @@ int main(int argc, char const *argv[]) {
     int K = 3;
     float R_incdec = 0.05;
     int R_lower = 18;
-    int R_scale = 5;
+    int R_scale = 2;
     float T_dec = 0.05;
-    int T_inc = 1;
+    int T_inc = 5;
     int T_lower = 2;
     int T_upper = 200;
-    int alpha = 10;
+    int alpha = 20;
     if(argc == 12) {
         N = stoi(argv[2]);
         K = stoi(argv[3]);

### 2019-02-18 paper/Paper.tex
a

--- 2019-02-16 paper/Paper.tex
@@
 
+\subsection{Post-processing}
+To enhance the performance of the algorithm is applied on the final mask a median-filter using a square kernel.
+Since the value of the pixels can be 0 or 255 is equivalent to perform a majority voting among the kernel's pixels to decide the value of the middle one.
+That has shown to reduce the noise on the final mask and in the meanwhile generate more connected components.
+
 \subsection{parameter settings}
@@
 decision threshold. Lowering it from 5 to 2 has shown to let the algorithm to recognize better the foreground object introducing very low noise.
-\item $T_{dec} = 0.05:$ If $x_i$ is background, $T_{dec}$ is the rate at which $T(x_i)$ is decreased.
+\item $T_{dec} = 0.05:$ If $x_i$ is background, $T_{dec}$ is the rate at which $T(x_i)$ is decreased. Decreasing it too fast can lead to include in the background model parts of foreground objects.
 \item $T_{inc} = 5:$ If $x_i$ is foreground, $T_{inc}$ is the rate at which $T(x_i)$ is increased.
@@
 Increasing it to 5 from the proposed value of 2 reduces faster the probability to put a steady foreground object in to one of the N components of the background model.
-\item $T_{lower}= 2:$ Lower bound of $T(x_i)$.
-\item $T_{upper}= 200:$ Upper bound of $T(x_i)$.
+\item $T_{lower}= 2:$ Lower bound of $T(x_i)$. It has been set to the value chosen in the original paper.
+\item $T_{upper}= 200:$ Upper bound of $T(x_i)$. It has been set to the value chosen in the original paper. 
 \end{itemize}

### 2019-02-18 paper/Paper.tex

--- 2019-02-17 paper/Paper.tex
@@
 Since one of the objective of the work is to have a robust classification of the foreground objects, we want to find the best values for the T parameters such that an object if classified as foreground stay as it for the longest period of time.
-INSERIRE IMMAGINI
+\begin{figure}[!t]
+    \centering
+    \subfloat[][$T_{inc}$ = 1]{
+        \includegraphics[width=0.45\textwidth]{Figures/T_inc_1.png}
+    }
+    \newline
+    \subfloat[][$T_{inc}$ = 5]{
+        \includegraphics[width=0.45\textwidth]{Figures/T_inc_5.png}
+    }
+    \caption{Have an higher $T_{inc}$ means decreasing faster the probability that a pixel is included in the background model once it is classified as foreground.
+Clearly visible in the two image reporting a car stopped for $30s$.}
+\end{figure}
 
@@
 
+\subsection{Post-processing}
+To enhance the performance of the algorithm is applied on the final mask a median-filter using a square kernel.
+Since the value of the pixels can be 0 or 255 is equivalent to perform a majority voting among the kernel's pixels to decide the value of the middle one.
+That has shown to reduce the noise on the final mask and in the meanwhile generate more connected components.
+The size chosen for the kernel after different experiments is $5\times 5$.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{Figures/median_blur.png}
+\caption{Comparison between output whitout post-processing 			and with median-blur post-processing}
+\end{figure}
+
 \subsection{Optimal Setting}

### 2019-02-19 paper/Paper.tex
Small changes

--- 2019-02-17 paper/Paper.tex
@@
 \section*{Proposed approach}
-To solve the presented problem we have studied deeply the PBAS algorithm, how it works, and how the hyperparameters could change the behaviour of it.
-After the studying of the algorithm has been chosen the parameters that have achived the better performance in our scenario.
-At the end have been studied how to enhanche the performance of the algorithm solving the issue of the presence of the shadows in the video, using techniques of shadows removal.
+Initially, we studied deeply the PBAS algorithm as originally thought: how it works and how the hyperparameters change its behaviour.
+Then we implemented it in a very efficient way using C++ and we further analyzed the quality of this first basic version.
+This shed light upon some improvements for important aspects as we see in the following sections.
+At the end, we have chosen the parameters that achieved the best performance in our scenario.
 \subsection{PBAS algorithm} \label{subsect:PBAS}

### 2019-02-19 paper/Paper.tex
Optical flow section. Corrections.

@@
 The value has been choosen considering the previous tradeoff.
-\item $K = 3:$ The number of components that have to be closer than $R(x_i)$ in order to set the pixel to be background. It has been left to the value proposed in the original paper.
+\item $K = 3:$ The number of components that have to be closer than $R(x_i)$ in order to set the pixel to be background.
+We raise it to 3 (instead of 2) because we want to be stricter in the background classification, since cars are often segmented only near the borders and masks present holes in the central area, probably because the car color is matched with some previous background pixels.
 \item $R_{incdec} = 0.05:$ Proportial constant of the controller for the variable $R(x_i)$.
@@
 
-\subsection{RGB vs Grey}
+\subsection{Long Period Experiments}
+
+\subsection{Algorithm Validation}
+
+\section*{Improvements}
+
+\subsection*{RGB vs Grey}
 We have experimented with a version of the algorithm working just on grey scale images and with one exploiting also color informations.
 The grey one is implementing exactly all the steps described in \ref{subsect:PBAS}.
-The RGB version is running the Grey version for any of the three channels and then the final mask is computed as the Logical-OR of the three masks obtained for each of the three channels.
-Experiments show that the RGB version has a better cars detection capabilities, but it proves to be computationally heavy, breaking totally the real time capabilities that we manage to achieve.
-Moreover, the behaviour proves to be better, but it doesn't justify the cost of having a non-realtime algorithm.
+The RGB version runs the greyscale version for each of the three channels in parallel threads and then the final mask is computed as the logical-OR of the three masks obtained from the three channels.
+Experiments show that the RGB version has better cars detection capabilities, but obviuosly computationally heavier.
+We could not be able to try with a 4-core CPU, but in theory this can bring times down to the grayscale version, since the 3 channels can be computed quickly expoliting parallelism.
 
@@
 
-\subsection{Long Period Experiments}
+\subsection*{Optical flow analysis}
+The algorithm is fully based on the variations of a pixel intensity over time.
+This cannot be enough in some situations to provide a correct segmentation, for example, we noticed that a traffic light is classified as a foreground element because it emits 3 different light colors that vary a lot from each other and very quickly from frame to frame (this problem has a more evident impact especially when using 3-channels videos).
+We can resort to optical flow to detect this kind of issues: in fact, optical flow provides a measure of the motion of pixels in consecutive frames, so in the previous case the flow of pixels in the area of the traffic light will not be present.
+So, we can discriminate pixels that have flow and pixels that have not, changing different parameters of the PBAS for these two categories.
 
-\subsection{Algorithm Validation}
+We built a model to determine which pixels had an "active" flow in the past, and in some sense these pixels represents road areas subject to traffic.
+We adopted the Farneback dense optical flow algorithm to get the flow of each pixel of the frame.
+Then, we checked which pixels had a flow magnitude over a certain threshold $W_{min}$ and we stored this information in a history of L previous samples.
+If the magnitude of the flow was greater then $W_{min}$ for L consecutive frames, than the pixel is classified as "flow-active".
+For all the "flow-active" pixels, we applied some ad-hoc adjustments for the PBAS parameters, in particular we decremented the update frequency T by a constant to penalyze the chances of those pixels to be inserted in the background model.
+Further changes can be done, however are out of the scope of this project.
 
-\subsection*{Conclusion}
+\begin{figure}[!t]
+    \centering
+    \subfloat[][No optical flow]{
+        \includegraphics[width=0.45\textwidth]{Figures/noflow.jpg}
+    }
+    \newline
+    \subfloat[][Optical flow]{
+        \includegraphics[width=0.45\textwidth]{Figures/flow.jpg}
+    }
+	\caption{Optical flow version compared with the naive version.
+The former one is aware of areas that have an "active flow" and put those pixels in the background model less frequent, e.g.
+the traffic light on the right.}
+\end{figure}
+
+\subsection*{Improvements and computational impact}
+We report here the time (in ms) required to analyze one frame using our implementation of the PBAS algorithm, with and without the proposed improvements.
+The algorithm has been run on a Macbook Pro mounting an Intel Core i5 2.3 GHz and 8 GB 2133MHz of LPDDR3 RAM, using as input 426x240 frames.
+Performances can drastically be raised by running it on a 4-core CPU (that is able to handle the 3 threads of the RBG channels in parallel) or on a GPU.
+
+\begin{table}[]
+\begin{tabular}{l|c}
+    Algorithm                                     & Average time \\
+    \hline
+    PBAS on grayscale video                       & 90ms         \\
+    PBAS on RBG video                             & 180ms        \\
+    PBAS on RBG video with optical flow analysis  & 200ms        
+\end{tabular}
+\end{table}
+
+\section*{Conclusion}
 We have provided a background subtraction system based on the PBAS algorithm.

### 2019-02-20 paper/Paper.tex
long period experiments

@@
 \subsection{Long Period Experiments}
+Since our goal is traffic monitoring 24/7, it is important to ensure that the algorithm can adapt as time passes to different light conditions, environmental changes, and traffic intensity variations.
+In order to do this, we let the algorithm run on ninty minutes long videos recording smooth light transitions (from day to night and viceversa), rough changes of environmental conditions and variations of traffic intensity.
+Experiments show that the algorithm adapts without any problem in those scenarios.
+This makes us believe that also in a real 24/7 scenario there wouldn't be any problem of adaptation.
+However, this behaviour is more than expected: looking at how the PBAS works, we can notice that the state, that is the values of the variables, would change in the same way during the starting iterations as it would after one day of non-stop run.
+
+\begin{figure}[!t]
+    \centering
+    \subfloat[][00:01:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/snow_60.jpg}
+    }
+    \newline
+    \subfloat[][00:26:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/snow_1561.jpg}
+    }
+    \newline
+    \subfloat[][00:35:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/snow_1921.jpg}
+    }
+    \newline
+    \subfloat[][01:04:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/snow_3850.jpg}
+    }
+    \newline
+    \subfloat[][01:29:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/snow_5377.jpg}
+    }
+    \caption{It is shown how the algorithm adapts to change of enviromental conditions. In this case it starts snowing, therefore the road gets dirtier than it was when the video started.}
+\end{figure}
 
@@
     }
-	\caption{RGB version compared with the Grey one. As we see, there is some difference present but it is negligible}
+	\caption{RGB version compared with the Grey one. As we see, there is some difference in the detection accuracy}
 \end{figure}
@@
 We have validated our work by visual inspection in particular and challenging situations, with long videos with environmental and traffic conditions changing over time.
-We further experimented with shadow removal methods and we tried to exploit motion informations.
+We further experimented with noise removal methods and we tried to exploit motion informations.
 Future work can be done in this direction, until arriving to a precise classification of noise and errors in the mask detected.

### 2019-02-20 paper/Paper.tex
little changes

--- 2019-02-19 paper/Paper.tex
@@
 \section*{Proposed approach}
-Initially, we studied deeply the PBAS algorithm as originally thought: how it works and how the hyperparameters change its behaviour.
-Then we implemented it in a very efficient way using C++ and we further analyzed the quality of this first basic version.
-This shed light upon some improvements for important aspects as we see in the following sections.
-At the end, we have chosen the parameters that achieved the best performance in our scenario.
+To solve the presented problem we have studied deeply the PBAS algorithm, how it works, and how the hyperparameters could change the behaviour of it.
+After the studying of the algorithm has been chosen the parameters that have achived the better performance in our scenario.
+At the end have been studied how to enhanche the performance of the algorithm solving the issue of the presence of the shadows in the video, using techniques of shadows removal.
 \subsection{PBAS algorithm} \label{subsect:PBAS}
@@
 zero too fast. \newline
-In the other ideal case of a fully static background, a classification as foreground is quite solid. \newline
+In the other ideal case of a fully static background, a classification as foreground is quite solid so $T(x_i)$ can be rapidly increased or decreased. \newline
 $T_{upper}, T_{lower}:$ control respectively the lowest and the highest probability of updating a pixel and putting it in the

### 2019-02-21 paper/Paper.tex
sunset long period vid

--- 2019-02-20 paper/Paper.tex
@@
     \centering
+    \subfloat[][00:00:30]{
+        \includegraphics[width=0.45\textwidth]{Figures/sunset_1.jpg}
+    }
+    \newline
+    \subfloat[][00:30:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/sunset_2.jpg}
+    }
+    \newline
+    \subfloat[][00:45:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/sunset_3.jpg}
+    }
+    \newline
+    \subfloat[][01:10:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/sunset_4.jpg}
+    }
+    \newline
+    \subfloat[][01:29:30]{
+        \includegraphics[width=0.45\textwidth]{Figures/sunset_5.jpg}
+    }
+    \caption{It is shown how the algorithm adapts to change of lighting conditions, in particular when the sun sets.}
+\end{figure}
+
+\begin{figure}[!t]
+    \centering
     \subfloat[][00:01:00]{
@@
     }
-	\caption{RGB version compared with the Grey one. As we see, there is some difference in the detection accuracy}
+	\caption{RGB version compared with the Grey one. As we see, there is little difference in the detection accuracy}
 \end{figure}

### 2019-02-21 paper/Paper.tex
24h description paper

--- 2019-02-20 paper/Paper.tex
@@
 
+\begin{figure}[!t]
+    \centering
+    \subfloat[][00:06:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_1.jpg}
+    }
+    \newline
+    \subfloat[][00:28:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_2.jpg}
+    }
+    \newline
+    \subfloat[][00:45:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_3.jpg}
+    }
+    \newline
+    \subfloat[][01:00:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_4.jpg}
+    }
+    \newline
+    \subfloat[][01:23:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_5.jpg}
+    }
+    \caption{It is shown how the algorithm adapts to change of lightining and traffic conditions during a 1h and 30 min video sample.}
+\end{figure}
+
 \subsection{Algorithm Validation}
@@
     PBAS on RBG video                             & 180ms        \\
-    PBAS on RBG video with optical flow analysis  & 200ms        
+    PBAS on RBG video with optical flow   & 200ms        
 \end{tabular}

### 2019-02-21 paper/Paper.tex
Added a figure for rbg vs grayscale version. Small graphical changes.

--- 2019-02-21 paper/Paper.tex
@@
 \begin{abstract}
-\textit{	Abstract-} In this paper we present a method for background subtraction with the aim to work on a 24/24h videos
-scenario, real time, robust to weather changes and capable to keep foreground objects detected for a
-large amount of time. The work is the base to implement a monitoring system for dangerous event in
-the road (possible scenario is the monitoring system on highways). The system can be expanded to
-detect events such as car driving in the wrong direction, car accidents or people crossing the road.
-The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a parametric background
-modeling paradigm, thus the background is modeled by a history of recently observed pixel values.
-The foreground decision depends on a decision threshold. The background update is based on a learning
-parameter. Both parameters are extended to dynamic per-pixel state variables and introduce dynamic
-controllers for each of them. Furthermore, both controllers are steered by an estimate of the background
-dynamics. All the hyperparameters of the models have been studied and tuned minutely to accomplish the
-aimed task.
+In this paper we present a method for background subtraction with the aim to work on a 24/24h videos scenario, real time, robust to weather changes and capable to keep foreground objects detected for a large amount of time.
+The work is the base to implement a monitoring system for dangerous event in the road (possible scenario is the monitoring system on highways).
+The system can be expanded to detect events such as car driving in the wrong direction, car accidents or people crossing the road.
+The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a parametric background modeling paradigm, thus the background is modeled by a history of recently observed pixel values.
+The foreground decision depends on a decision threshold.
+The background update is based on a learning parameter.
+Both parameters are extended to dynamic per-pixel state variables and introduce dynamic controllers for each of them.
+Furthermore, both controllers are steered by an estimate of the background dynamics.
+All the hyperparameters of the models have been studied and tuned minutely to accomplish the aimed task.
+Then we will also explain which additions we introduce in the algorithm to improve its performance in our particular scenario.
 \end {abstract}
@@
 of a video and saying if that pixels are part of background or foreground.
-At the end, the algorithm take as input a video and output a binary mask, where 1s are foreground pixels and 0 background pixels.
+At the end, the algorithm take as input a video and output a binary mask, where 1s are foreground pixels and 0s background pixels.
 We restricted our domain to traffic monitoring, in particular this work should put the basis for an automated
@@
 allow the system to work in real-time. In \cite{pbas_and_scene_analysis_fpga}, we can find an integration of
-the basic algorithm in a FPGA device, which exploit and hardware implementation to reduce a lot the
-computational time.
+the basic algorithm in a FPGA device, which exploit and hardware implementation to reduce a lot the computational time.
 
@@
 Over the recent past, a multitude of algorithms and methods for background modeling have been developed.
-One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM)
-
-\cite{gmm}. Here, each pixels is modeled as a mixture of weighted Gaussian distributions. Pixels, which
-are detected as background are used to improve the Gaussian mixtures by an iterative update rule. A very
-important non-parametric method is the ViBe \cite{vibe}. Each pixel in the background model is defined by
-a history of the N most recent image values at each pixel and uses a random scheme to update them.
-More over, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another
-random selection method. The preceding scheme is very similar to the approach followed by the PBAS algorithm.
-It can be categorized as a non-parametric method, since it uses a history of N image values as the background
-model, and uses a random update rule similar to the one used by the ViBe algorithm. However, in ViBe, the
-randomness parameters as well as the decision threshold are fixed for all pixels. In contrast, in the PBAS
-algorithm these values are not treated as parameters, but instead as adaptive state variables, which can
-dynamically change over time for each pixel separately.
+One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM) \cite{gmm}.
+Here, each pixels is modeled as a mixture of weighted Gaussian distributions.
+Pixels, which are detected as background are used to improve the Gaussian mixtures by an iterative update rule.
+A very important non-parametric method is the ViBe \cite{vibe}.
+Each pixel in the background model is defined by a history of the N most recent image values at each pixel and uses a random scheme to update them.
+More over, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another random selection method.
+The preceding scheme is very similar to the approach followed by the PBAS algorithm.
+It can be categorized as a non-parametric method, since it uses a history of N image values as the background model, and uses a random update rule similar to the one used by the ViBe algorithm.
+However, in ViBe, the randomness parameters as well as the decision threshold are fixed for all pixels.
+In contrast, in the PBAS algorithm these values are not treated as parameters, but instead as adaptive state variables, which can dynamically change over time for each pixel separately.
 
@@
 At the end, we have chosen the parameters that achieved the best performance in our scenario.
+
 \subsection{PBAS algorithm} \label{subsect:PBAS}
 
+This section describes the Pixel-Based Adaptive Segmenter, which follows a non-parametric paradigm.
+Thus, every pixel $x_i$ is modeled by an array of recently observed background values.
+The method consists of several components which are depicted as a state machine in Figure 1.
+As a central component, the decision block decides for or against foreground based on the current image and a background model $B(x_i)$.
+This decision is based on the per-pixel threshold $R(x_i)$.
+Moreover, the background model has to be updated over time in order to allow for gradual background changes.
+In the model, this update depends on a per-pixel learning parameter $T(x_i)$.
+\newline
 \begin{figure*}
@@
 
-This section describes the Pixel-Based Adaptive Segmenter, which follows a non-parametric paradigm. Thus,
-every pixel $x_i$ is modeled by an array of recently observed background values.
-The method consists of several components which are depicted as a state machine in Figure 1.
-As a central component, the decision block decides for or against foreground
-based on the current image and a background model $B(x_i)$.
-This decision is based on the per-pixel threshold $R(x_i)$.
-Moreover, the background model has to be updated over time in order to allow for gradual background changes.
-In the model, this update depends on a per-pixel learning parameter $T(x_i)$.
-\newline The background model $B(x_i)$ is defined by an array of N recently observed pixel values:
+The background model $B(x_i)$ is defined by an array of N recently observed pixel values:
 \begin{equation}
@@
 In the other ideal case of a fully static background, a classification as foreground is quite solid so $T(x_i)$ can be rapidly increased or decreased. \newline
-$T_{upper}, T_{lower}:$ control respectively the lowest and the highest probability of updating a pixel and putting it in the
-background model $B$. \newline
-$T_{inc}, T_{dec}:$ are fixed parameters and control the update of the controller T of each pixel, has been chosen to differentiate the two quantity since the pixels that are background are usually the largest part so choosing different parameters can lead to a better perfomance.\newline
+$T_{upper}$, $T_{lower}$: control respectively the lowest and the highest probability of updating a pixel and putting it in the background model $B$. \newline
+$T_{inc}$, $T_{dec}$: are fixed parameters and control the update of the controller T of each pixel, has been chosen to differentiate the two quantity since the pixels that are background are usually the largest part so choosing different parameters can lead to a better perfomance.\newline
 Since one of the objective of the work is to have a robust classification of the foreground objects, we want to find the best values for the T parameters such that an object if classified as foreground stay as it for the longest period of time.
@@
 
-\subsection*{$\alpha$}
+\subsection*{$\bm{\alpha}$}
 From eq \ref{eq:dist} we see that $\alpha$ tells how much to weight the image gradients
@@
     \centering
-    \subfloat[][00:00:30]{
-        \includegraphics[width=0.45\textwidth]{Figures/sunset_1.jpg}
-    }
-    \newline
-    \subfloat[][00:30:00]{
-        \includegraphics[width=0.45\textwidth]{Figures/sunset_2.jpg}
-    }
-    \newline
-    \subfloat[][00:45:00]{
-        \includegraphics[width=0.45\textwidth]{Figures/sunset_3.jpg}
-    }
-    \newline
-    \subfloat[][01:10:00]{
-        \includegraphics[width=0.45\textwidth]{Figures/sunset_4.jpg}
-    }
-    \newline
-    \subfloat[][01:29:30]{
-        \includegraphics[width=0.45\textwidth]{Figures/sunset_5.jpg}
-    }
-    \caption{It is shown how the algorithm adapts to change of lighting conditions, in particular when the sun sets.}
-\end{figure}
-
-\begin{figure}[!t]
-    \centering
     \subfloat[][00:01:00]{
@@
     }
-	\caption{RGB version compared with the Grey one. As we see, there is little difference in the detection accuracy}
+    \newline
+    \subfloat[][Gray and RBG on intermittent motion video]{
+        \includegraphics[width=0.45\textwidth]{Figures/rgbvsgray.jpg}
+    }
+    \caption{Comparison between grayscale and RBG versions of the PBAS.
+As we see, there are some differences in the detection accuracy.
+In good light situations, e.g.
+sunny days, the RGB version provide better foreground masks and less noise.}
 \end{figure}

### 2019-02-22 eval.cpp
eval fix

@@ -66,9 +66,8 @@ Metrics eval(string path) {
     for (int i=0; i < inputs.size(); i++){
         Mat input, gt, mask;
         input = imread(inputs[i].string(), CV_LOAD_IMAGE_COLOR);
-
-        cvtColor(input, input, cv::COLOR_RGB2GRAY);
-        mask = *(pbas->process(input));
+        // cvtColor(input, input, cv::COLOR_RGB2GRAY);
+        mask = *(pbas->process(&input));
 
         if (i+1 >= temporal_ROI_beg && i+1 <= temporal_ROI_end) {
             gt = imread(gts[i].string(), CV_8UC1);
@@ -94,8 +93,8 @@ Metrics eval(string path) {
     float fnr = (float)fn/(tn+fp);
     float pwc = 100*((float)fn+fp)/(tp+fn+fp+tn);
 
-    // cout << "tp: " + to_string(tp) + " -- tn: " + to_string(tn) + " -- fn: " + to_string(fn) + " -- fp: " + to_string(fp) + "\n" << endl;
-    // cout << "re: " + to_string(re) + " -- sp: " + to_string(sp) + " -- fpr: " + to_string(fpr) + " -- fnr: " + to_string(fnr) + " -- pwc: " + to_string(pwc) + " -- f1: " + to_string(f1) + " -- pr: " + to_string(pr) << endl;
+    cout << "tp: " + to_string(tp) + " -- tn: " + to_string(tn) + " -- fn: " + to_string(fn) + " -- fp: " + to_string(fp) + "\n" << endl;
+    cout << "re: " + to_string(re) + " -- sp: " + to_string(sp) + " -- fpr: " + to_string(fpr) + " -- fnr: " + to_string(fnr) + " -- pwc: " + to_string(pwc) + " -- f1: " + to_string(f1) + " -- pr: " + to_string(pr) << endl;
     
     return {re, sp, fpr, fnr, pwc, f1, pr};
 }

### 2019-02-22 paper/Paper.tex

@@
     \caption{It is shown how the algorithm adapts to change of enviromental conditions. In this case it starts snowing, therefore the road gets dirtier than it was when the video started.}
+\end{figure}
+
+\begin{figure}[!t]
+    \centering
+    \subfloat[][00:06:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_1.jpg}
+    }
+    \newline
+    \subfloat[][00:28:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_2.jpg}
+    }
+    \newline
+    \subfloat[][00:45:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_3.jpg}
+    }
+    \newline
+    \subfloat[][01:00:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_4.jpg}
+    }
+    \newline
+    \subfloat[][01:23:00]{
+        \includegraphics[width=0.45\textwidth]{Figures/24h_5.jpg}
+    }
+    \caption{It is shown how the algorithm adapts to change of lightining and traffic conditions during a 1h and 30 min video sample.}
 \end{figure}

### 2019-02-22 paper/Paper.tex
fixes on paper for final version

@@
 The system can be expanded to detect events such as car driving in the wrong direction, car accidents or people crossing the road.
-The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a parametric background modeling paradigm, thus the background is modeled by a history of recently observed pixel values.
+The model studied is the PBAS (Pixel-Based Adaptive Segmenter): it follows a non-parametric background modeling paradigm, thus the background is modeled by a history of recently observed pixel values.
 The foreground decision depends on a decision threshold.
@@
 for example, surveillance systems, autonomous vehicles, semantic image analysis. Background subtraction
-consists in finding a category for every pixel in a single or (like in this case) in a frames sequence
-of a video and saying if that pixels are part of background or foreground.
+consists in classifying as foreground or background every single pixel belonging to a frame of a video sequence.
 At the end, the algorithm take as input a video and output a binary mask, where 1s are foreground pixels and 0s background pixels.
@@
 a lot from the day and our algorithm has to adapt continuosly due to this facts. 
-Other problems that we faced are shadows and intermittent object motion.
-From the majority of the current algorithms, shadows are treated as foreground because they move along with
-objects; a little number of them adopts a 3-class classification (background, foreground, shadow). We
-decided to provide a binary classification, considering shadows as part of the background. With regards to
-the latter problem, here the challenge is to correctly detect an initially stationary object that begins to
+Another problem that we faced is intermittent object motion. Here the challenge is to correctly detect an initially stationary object that begins to
 move or when an object that was static starts moving again. We fine-tuned some parameters to make sure
@@
 Each pixel in the background model is defined by a history of the N most recent image values at each pixel and uses a random scheme to update them.
-More over, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another random selection method.
-The preceding scheme is very similar to the approach followed by the PBAS algorithm.
+Moreover, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another random selection method.
+The preceding scheme is very similar to the approach followed by the PBAS \cite{pbas} algorithm.
 It can be categorized as a non-parametric method, since it uses a history of N image values as the background model, and uses a random update rule similar to the one used by the ViBe algorithm.
@@
 $F = 1$ implies foreground.
-The decision involves two parameters: the distance threshold $R(x_i)$,which is defined for each pixel separately and which can change dynamically; and the minimum number $K$, which is a fixed global parameter.
+The decision involves two parameters: the distance threshold $R(x_i)$, which is defined for each pixel separately and which can change dynamically; and the minimum number $K$, which is a fixed global parameter.
 \paragraph{Update B}
@@
 Whenever an update of $B_k(x_i)$ is carried out, the currently observed minimal distance:
-\[ d_{min}(x_i) = min_k dist(I(x_i), B_k(x_i)) \] is written to this array: $D_k(x_i) \leftarrow d_{min}(x_i)$.
+\[ d_{min}(x_i) = min_k \{dist(I(x_i), B_k(x_i))\} \] is written to this array: $D_k(x_i) \leftarrow d_{min}(x_i)$.
 Thus, a history of minimal decision distances is created.
@@
 Then, we show some real-case usages of the algorithm, comparing different environmental conditions that we found in 24/24h live videos.
-Next, are reported experiments made with
 
@@
 The greater, the more complex backgrounds can be handled but at the cost of heavier
-computation. At $N=35$, experimental results show that performance saturates.
+computation.
 \begin{figure}[!t]
@@
 It is slowly increased when the pixel is classified as background and slowly decreased when the pixel is classified as foreground.
-That leads to have a better foreground detection for multiple reasons.
-\newline
-In case of highly dynamic background (i.e.
+That leads to have a better foreground detection for multiple reasons: in case of highly dynamic background (i.e.
 big $\overline{d}_{min}(x_i)$), the learning parameter $T(x_i)$ stays constant or only slightly changes.
-In this case of highly dynamic background, erroneously detected foreground will not remain for long, be-
-cause the probability to update $p = 1/T(x_i)$ does not reach
+In this case of highly dynamic background, erroneously detected foreground will not remain for long, because the probability to update $p = 1/T(x_i)$ does not reach
 zero too fast. \newline
@@
 $T_{upper}$, $T_{lower}$: control respectively the lowest and the highest probability of updating a pixel and putting it in the background model $B$. \newline
-$T_{inc}$, $T_{dec}$: are fixed parameters and control the update of the controller T of each pixel, has been chosen to differentiate the two quantity since the pixels that are background are usually the largest part so choosing different parameters can lead to a better perfomance.\newline
+$T_{inc}$, $T_{dec}$: are fixed parameters and control the update of the controller T of each pixel.\newline
 Since one of the objective of the work is to have a robust classification of the foreground objects, we want to find the best values for the T parameters such that an object if classified as foreground stay as it for the longest period of time.
@@
 \subsection*{R}
-Eq \ref{eq:R} is similar to the equation of a pure proportional controller.
+Eq (\ref{eq:R}) is similar to the equation of a pure proportional controller.
 In this case the $R(x,y)$ has to follow $d_{minavg}(x,y)R_{scale}$. So:
 \begin{itemize}
-    \item $R_{incdec}$: it can be seen as the propositional constant of the controller: setting too high lead to oscillation around the target and if set too high leads to divergence.
+    \item $R_{incdec}$: it can be seen as the propositional constant of the controller: setting too high can lead to oscillation around the target and even to divergence.
 Setting too low leads to slow convergence to the target values, making the algorithm too slow to adapt to dynamic backgrounds.
@@
     but also decreases true positives; setting too low, will increase false positives.
-    \item $R_{lower}$: sets a limit on how much $R$ can decrease its value. Setting too
-    low, it will lead to a lot of false positives in area that are foreground,
-    especially if the video is noise, so especially is zones that are clearly
-    background are changing their values of high amounts due to noise. Setting too
-    high it will reduce true positives. So the choice of it should be lead by the
+    \item $R_{lower}$: sets a limit on how much $R$ can decrease its value. Setting it too
+    low will lead to a lot of false positives,
+   especially in zones that are changing their values of high amounts due to noise. Setting too
+    high it will reduce true positives. So its value should be led by the
     video quality, more than the application domain.
@@
 \subsection*{$\bm{\alpha}$}
-From eq \ref{eq:dist} we see that $\alpha$ tells how much to weight the image gradients
+From eq (\ref{eq:dist}) we see that $\alpha$ tells how much to weight the image gradients
 with respect to the image values in the distance computation.
@@
 %\end{figure}
-As we see, the image gradients give a good cue to distinguish the street from the cars
+The image gradients give a good cue to distinguish the street from the cars
 (image absolute values are not, since many cars have almost the same color of the
@@
 Since the value of the pixels can be 0 or 255 is equivalent to perform a majority voting among the kernel's pixels to decide the value of the middle one.
-That has shown to reduce the noise on the final mask and in the meanwhile generate more connected components.
+That has shown to reduce the noise on the final mask and in the meanwhile generate more connected components (Figure \ref{fig:median}).
 The size chosen for the kernel after different experiments is $5\times 5$.
@@
 \caption{Comparison between output whitout post-processing and with median-blur post-processing}
+\label{fig:median}
 \end{figure}
@@
 \subsection{Optimal Setting}
-The proposed method has a multitude of tunable parameters, their values have been chosen focusing the attention on having an algorithm robust to changing on weather condition and capable of identifying foreground object even if they stay steady for a period of time (i.e.
+The proposed method has a multitude of tunable parameters, their values have been chosen focusing the attention on having an algorithm robust to changing on weather condition and capable of identifying foreground object even if they stay steady for a period of time (e.g.
 cars stopped to a traffic light), still guaranteeing a real time behaviour.
@@
 It has been left to the value proposed in the original paper, since we found that raising it would lead to a significant decrease in true positive and lowering it would let too much noise in.
-\item $R_{scale}=2$ Scaling factor in the controller for the decision threshold.
-Setting it to 2 has shown to let the algorithm to recognize better the foreground object while introducing very low noise.
+\item $R_{scale}=2:$ Scaling factor in the controller for the decision threshold.
+Setting it to 2 has shown to let the algorithm recognize better the foreground object while introducing very low noise.
 \item $T_{dec} = 0.05:$ If $x_i$ is background, $T_{dec}$ is the rate at which $T(x_i)$ is decreased.
@@
 
-\subsection{Algorithm Validation}
-
 \section*{Improvements}
@@
 The RGB version runs the greyscale version for each of the three channels in parallel threads and then the final mask is computed as the logical-OR of the three masks obtained from the three channels.
-Experiments show that the RGB version has better cars detection capabilities, but obviuosly computationally heavier.
+Experiments show that the RGB version has slightly better cars detection capabilities, but obviuosly computationally heavier.
 We could not be able to try with a 4-core CPU, but in theory this can bring times down to the grayscale version, since the 3 channels can be computed quickly expoliting parallelism.

### 2019-02-22 paper/Paper.tex
Small corrections

@@
 consists in classifying as foreground or background every single pixel belonging to a frame of a video sequence.
-At the end, the algorithm take as input a video and output a binary mask, where 1s are foreground pixels and 0s background pixels.
+In synthesis, the algorithm take as input a video and output a binary mask, where 1s are foreground pixels and 0s background pixels.
 We restricted our domain to traffic monitoring, in particular this work should put the basis for an automated
@@
 Over the recent past, a multitude of algorithms and methods for background modeling have been developed.
-One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM) \cite{gmm}.
-Here, each pixels is modeled as a mixture of weighted Gaussian distributions.
-Pixels, which are detected as background are used to improve the Gaussian mixtures by an iterative update rule.
-A very important non-parametric method is the ViBe \cite{vibe}.
-Each pixel in the background model is defined by a history of the N most recent image values at each pixel and uses a random scheme to update them.
+One of the most prominent and most widely used methods are those based on Gaussian Mixture Models (GMM) \cite{gmm}: each pixel is modelled as a mixture of weighted Gaussian distributions.
+Pixels which are detected as background are used to improve the Gaussian mixtures by an iterative update rule.
+Another important non-parametric method is the ViBe \cite{vibe}.
+Each pixel in the background model is defined by a history of the N most recent image pixel values and it uses a random scheme to update them.
 Moreover, updated pixels can ”diffuse” their current pixel value into neighboring pixel using another random selection method.
@@
 $F = 1$ implies foreground.
-The decision involves two parameters: the distance threshold $R(x_i)$, which is defined for each pixel separately and which can change dynamically; and the minimum number $K$, which is a fixed global parameter.
+The decision involves two parameters: the distance threshold $R(x_i)$, which is defined for each pixel separately and which can change dynamically, and the minimum number $K$, which is a fixed global parameter.
+The function $dist$ is a distance measure, as explained in equation (\ref{eq:dist}).
 \paragraph{Update B}
@@
 are currently background (i.e. $F(x_i) = 0$).
-Updating means that for a certain index $k \in \{1...N \}$ (chosen uniformly at random), the corresponding background model
-value $B_k(x_i)$ is replaced by the current pixel value $I(x_i)$.
-This allows the current pixel value to be ”learned” into
-the background model. This update, however, is only performed with probability $p = 1/T(x_i)$. Otherwise no update
-is carried out at all. Therefore, the parameter $T(x_i)$ defines
-the update rate. The higher $T(x_i)$ the less likely a pixel
-will be updated.
-It is also updated (with probability $p = 1/T(x_i)$) a randomly chosen neighboring pixel $y_i \in N(x_i)$. Thus, the
-background model $B_k(y_i)$ at this neighboring pixel is replaced by its current pixel value $V(y_i)$.
+Updating means that for a certain index $k \in \{1...N \}$ (chosen uniformly at random), the corresponding background model value $B_k(x_i)$ is replaced by the current pixel value $I(x_i)$.
+This allows the current pixel value to be ”learned” into the background model. This update, however, is only performed with probability $p = 1/T(x_i)$. Otherwise no update
+is carried out at all. Therefore, the parameter $T(x_i)$ defines the update rate. The higher $T(x_i)$ the less likely a pixel will be updated.
+It is also updated (with probability $p = 1/T(x_i)$) a randomly chosen neighboring pixel $y_i \in N(x_i)$.
+Thus, the background model $B_k(y_i)$ at this neighboring pixel is replaced by its current pixel value $V(y_i)$.
 \paragraph{Update R}
@@
 The average of these values $\overline{d}_{min}(x_i) = \frac{1}{N} \sum_k D_k(x_i)$ is a measure of the background dynamics.
-Thanks to that
-the decision threshold can be dynamically adapted as follows:
+Thanks to that, the decision threshold can be dynamically adapted as follows:
 \begin{equation} \label{eq:R}
@@
 \end{equation}
-Here, $R_{inc/dec}$, $R_{scale}$ are fixed parameters. This can be seen as a dynamic controller for the state variable $R(x_i)$.
+Here, $R_{inc/dec}$ and $R_{scale}$ are fixed parameters. This can be seen as a dynamic controller for the state variable $R(x_i)$.
 \paragraph{Update T}
@@
 $\overline{I_m}$ is the average gradient magnitude over the last observed frame.
-Thus, the fraction $\frac{\alpha}{\overline{I^m}}$ weighs the importance
-of pixel values against the gradient magnitude
+Thus, the fraction $\frac{\alpha}{\overline{I^m}}$ weighs the importance of pixel values against the gradient magnitude.
 
@@
 
-In this section we first report the results obtained by varying the main model parameters.
+In this section, we first report the results obtained by varying the main model parameters.
 We also include the adopted settings that perform best in our scenario.
@@
 $N$ represents how many matrices are kept in memory to model the background.
-Each of them stores the history of pixels values.
-The greater, the more complex backgrounds can be handled but at the cost of heavier
+Each of them stores the history of pixels values (so $N$ entries).
+The greater, the more complex backgrounds can be handled, but at the cost of heavier
 computation.
@@
 \subsection{Post-processing}
-To enhance the performance of the algorithm is applied on the final mask a median-filter using a square kernel.
-Since the value of the pixels can be 0 or 255 is equivalent to perform a majority voting among the kernel's pixels to decide the value of the middle one.
+To refine the performance of the algorithm, a median-filter is applied to the final mask using a square kernel.
+Since the values of the pixels can be 0 or 255 this is equivalent to perform a majority voting among the kernel's pixels to decide the value of the middle one.
 That has shown to reduce the noise on the final mask and in the meanwhile generate more connected components (Figure \ref{fig:median}).
@@
 \item $T_{upper}= 200:$ Upper bound of $T(x_i)$.
-\item $\alpha = 20:$ Weight to give to image gradients. The proposed value is high because this shows to be a good cue for distinguishing car objects
+\item $\alpha = 20:$ Weight to give to image gradients. The proposed value is high because this shows to be a good cue for distinguishing car objects.
 \end{itemize}
@@
 Since our goal is traffic monitoring 24/7, it is important to ensure that the algorithm can adapt as time passes to different light conditions, environmental changes, and traffic intensity variations.
-In order to do this, we let the algorithm run on ninty minutes long videos recording smooth light transitions (from day to night and viceversa), rough changes of environmental conditions and variations of traffic intensity.
+In order to do this, we let the algorithm run on ninety minutes long videos recording smooth light transitions (from day to night and viceversa), rough changes of environmental conditions and variations of traffic intensity.
 Experiments show that the algorithm adapts without any problem in those scenarios.
-This makes us believe that also in a real 24/7 scenario there wouldn't be any problem of adaptation.
-However, this behaviour is more than expected: looking at how the PBAS works, we can notice that the state, that is the values of the variables, would change in the same way during the starting iterations as it would after one day of non-stop run.
+This makes us believe that also in a real 24/7 scenario there wouldn't be any adaptation problems.
+However, this behaviour is better than expected: looking at how the PBAS works, we can notice that the state (that is the values of the variables) would change in the same way during the starting iterations as it would after one day of non-stop run.
 
@@
 We adopted the Farneback dense optical flow algorithm to get the flow of each pixel of the frame.
-Then, we checked which pixels had a flow magnitude over a certain threshold $W_{min}$ and we stored this information in a history of L previous samples.
+Then, we checked which pixels had a flow magnitude over a certain threshold $W_{min}$ and we stored this information in a history of $L$ previous samples.
 If the magnitude of the flow was greater then $W_{min}$ for L consecutive frames, than the pixel is classified as "flow-active".
-For all the "flow-active" pixels, we applied some ad-hoc adjustments for the PBAS parameters, in particular we decremented the update frequency T by a constant to penalyze the chances of those pixels to be inserted in the background model.
+For all the "flow-active" pixels, we applied some ad-hoc adjustments for the PBAS parameters, in particular we decremented the update frequency $T$ by a constant to penalyze the chances of those pixels to be inserted in the background model.
 Further changes can be done, however are out of the scope of this project.

### 2019-02-22 paper/Paper.tex
Added final revision of the paper. Updated readme and sources to run the algorithm.

@@
 \paragraph{Update R}
-To change the decision treshold $R(x_i)$ is saved an array of recently observed pixel values in the background model $B(x_i)$, we also create an array $\mathbf{D}(x_i) = \{D_1(x_i ),...,D_N(x_i)\}$ of minimal decision distances.
+To change the decision treshold $R(x_i)$ is saved an array of recently observed pixel values in the background model $B(x_i)$.
+We also create an array $\mathbf{D}(x_i) = \{D_1(x_i ),...,D_N(x_i)\}$ of minimal decision distances.
 Whenever an update of $B_k(x_i)$ is carried out, the currently observed minimal distance:
@@
 $T(x_i)$. To alleviate the problem, the idea is to introduce a
-(second) dynamic controller for $T(x_i)$ following the assumption that pixels are mostly wrongly classified as foreground in areas of high dynamic background the strength of the adjustment in the controller can be adapted using the dynamic estimator $\overline{d}_{min}(x_i)$.
+(second) dynamic controller for $T(x_i)$ following the assumption that pixels are mostly wrongly classified as foreground in areas of high dynamic background.
+The strength of the adjustment in the controller can be adapted using the dynamic estimator $\overline{d}_{min}(x_i)$.
 So can be defined:

### 2019-02-22 pbas.cpp

@@ -14,7 +14,7 @@ using namespace std;
 using namespace cv;
 
 
-PBAS::PBAS() {
+PBASgray::PBASgray() {
     N = 30;
     K = 3;
     R_incdec = 0.05;
@@ -33,7 +33,7 @@ PBAS::PBAS() {
     init();
 }
 
-PBAS::PBAS(int N, int K=2, float R_incdec=0.05, int R_lower=18, int R_scale=5, float T_dec=0.05, int T_inc=1, int T_lower=2, int T_upper=200, int alpha = 10)
+PBASgray::PBASgray(int N, int K=2, float R_incdec=0.05, int R_lower=18, int R_scale=5, float T_dec=0.05, int T_inc=1, int T_lower=2, int T_upper=200, int alpha = 10)
 {
     this->N = N;
     this->K = K;
@@ -49,7 +49,7 @@ PBAS::PBAS(int N, int K=2, float R_incdec=0.05, int R_lower=18, int R_scale=5, f
     init();
 }
 
-void PBAS::init() {
+void PBASgray::init() {
     //initialize random seed
     //srand (time(NULL));
 
@@ -64,7 +64,7 @@ void PBAS::init() {
     displacement_vec.push_back(make_pair(0, -1));
 }
 
-PBAS::~PBAS() {
+PBASgray::~PBASgray() {
     frame.release();
     frame_grad.release();
     median.release();
@@ -77,15 +77,15 @@ PBAS::~PBAS() {
     displacement_vec.clear();
 }
 
-float PBAS::distance(uint8_t a, uint8_t b) {
+float PBASgray::distance(uint8_t a, uint8_t b) {
     return abs(a-b);
 }
 
-float PBAS::distance(uint8_t p, uint8_t p_grad, uint8_t g, uint8_t g_grad) {
+float PBASgray::distance(uint8_t p, uint8_t p_grad, uint8_t g, uint8_t g_grad) {
     return (this->alpha/this->I_m) * abs((int)p_grad - (int)g_grad) + abs((int)p - (int)g); 
 }
 
-Mat PBAS::gradient_magnitude(Mat* frame){
+Mat PBASgray::gradient_magnitude(Mat* frame){
     Mat grad;
     int scale = 1;
     int delta = 0;
@@ -109,7 +109,7 @@ Mat PBAS::gradient_magnitude(Mat* frame){
     return grad;
 }
 
-void PBAS::init_Mat(Mat* matrix, float initial_value){
[…]

### 2019-02-22 pbas.h

@@ -4,7 +4,7 @@
 using namespace std;
 using namespace cv;
 
-class PBAS
+class PBASgray
 {
     private:
         int N;
@@ -73,9 +73,9 @@ class PBAS
         // shadows with paperino
         Mat shadows_corner(Mat* frame, Mat* mask);
     public:
-        PBAS();
-        PBAS(int N, int K, float R_incdec, int R_lower, int R_scale, float T_dec, int T_inc, int T_lower, int T_upper, int alpha);
-        ~PBAS();
+        PBASgray();
+        PBASgray(int N, int K, float R_incdec, int R_lower, int R_scale, float T_dec, int T_inc, int T_lower, int T_upper, int alpha);
+        ~PBASgray();
 
         Mat* process(const Mat* frame);
         void showCVMat(Mat matrix, bool normalize, string window_name);

### 2019-02-22 pbas_rbg.cpp

@@ -86,7 +86,6 @@ class PBAS
         Mat gradient_magnitude(const Mat &frame);
 
     public:
-        PBAS();
         PBAS(int N, int K, float R_incdec, int R_lower, int R_scale, float T_dec, int T_inc, int T_lower, int T_upper, int alpha);
         ~PBAS();
         
@@ -207,9 +206,6 @@ Mat PBAS::process(const Mat &frame) {
     medianBlur(final_mask,final_mask,3);
 
     frame.copyTo(estimatedBg, 255-final_mask);
-
-    Mat toshow; T[0].convertTo(toshow, CV_8UC1);
-    imshow("T", toshow); moveWindow("T", 0, 550);
 
     return final_mask;
 }

### 2019-02-22 run_PBASgrayscale.cpp

@@ -1,0 +1,46 @@
+#include <iostream>
+#include <string>
+#include <opencv2/opencv.hpp>
+#include <sys/stat.h> 
+#include "pbas.cpp"
+
+using namespace std;
+using namespace cv;
+
+int current_frame = 0;
+float FPS = 30;
+
+int main(int argc, char const *argv[]) {
+    if(argc != 2) {
+        cout << "Missing parameter: video path." << endl;
+        exit(0);
+    }
+    string videopath = argv[1];
+    VideoCapture cap(videopath);
+
+    PBASgray *pbas = new PBASgray(30);
+
+    Mat frame;
+    Mat *mask;
+    
+    while(1) {
+        cap >> frame;
+        if (frame.empty()) break;
+
+        mask = pbas->process(&frame);
+
+        imshow("Grayscale", frame);
+        if(current_frame == 0) moveWindow("Grayscale", 0,20);
+        imshow("Grayscale mask", *mask);
+        if(current_frame == 0) moveWindow("Grayscale mask", 425, 20);
+
+        char c=(char)waitKey(25);
+        if(c==27) break;
+
+        current_frame ++;
+    }
+    pbas->~PBASgray();
+    cap.release();
+    destroyAllWindows();
+    return 0;
+}

### 2019-02-22 run_PBASrbg.cpp

@@ -1,0 +1,47 @@
+#include <iostream>
+#include <string>
+#include <opencv2/opencv.hpp>
+#include <sys/stat.h> 
+#include "pbas_rbg.cpp"
+
+using namespace std;
+using namespace cv;
+
+int current_frame = 0;
+float FPS = 30;
+
+int main(int argc, char const *argv[]) {
+    if(argc != 2) {
+        cout << "Missing parameter: video path." << endl;
+        exit(0);
+    }
+    string videopath = argv[1];
+    VideoCapture cap(videopath);
+
+    PBAS *pbas = new PBAS(30);
+    pbas->verbose = false;
+
+    Mat frame;
+    Mat mask;
+    
+    while(1) {
+        cap >> frame;
+        if (frame.empty()) break;
+
+        mask = pbas->process(frame);
+
+        imshow("RBG", frame);
+        if(current_frame == 0) moveWindow("RBG", 0,20);
+        imshow("RBG mask", mask);
+        if(current_frame == 0) moveWindow("RBG mask", 425, 20);
+
+        char c=(char)waitKey(25);
+        if(c==27) break;
+
+        current_frame ++;
+    }
+    pbas->~PBAS();
+    cap.release();
+    destroyAllWindows();
+    return 0;
+}

### 2019-02-25 pbas.cpp
Removed visualization of the gradient

@@ -152,8 +152,8 @@ Mat* PBASgray::process(const Mat* frame) {
     this->frame_grad = gradient_magnitude(&this->frame);
     this->I_m = mean(this->frame_grad).val[0];
 
-    showCVMat(this->frame_grad, true, "grad");
-    moveWindow("grad", 900, 900);
+    //showCVMat(this->frame_grad, true, "grad");
+    //moveWindow("grad", 900, 900);
 
     // B, D, d_minavg initialization
     if (B.size() == 0) {
